---
title: "Design, power, and cleaning tools from three audit studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Design, power, and cleaning tools from three audit studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.2, fig.align = "center",
  warning = FALSE, message = FALSE
)
has_geepack  <- requireNamespace("geepack",  quietly = TRUE)
has_glmmTMB  <- requireNamespace("glmmTMB",  quietly = TRUE)
has_pwr      <- requireNamespace("pwr",      quietly = TRUE)
has_fastLink <- requireNamespace("fastLink", quietly = TRUE)
```

Three real mystery-caller studies each hand-rolled analysis code that belonged
in the package. This vignette walks the tools generalized from them, grouped by
where they came from:

- a **matched-pair** audit (the same practice called under several caller
  personas) — design-integrity checks, a population-average sensitivity model,
  and two paired figures;
- an **ENT geographic-access power** study — power calculators the package's
  count/paired tools did not cover;
- the original **urogynecology** study — raw-field cleaning for messy call logs.

```{r}
library(mysterycall)
```

# 1. Design integrity for a matched multi-scenario audit

A matched design only draws information from practices reached under *both* arms
of a contrast, so before analysing anything it is worth knowing how complete the
design actually is — and whether a mistyped practice name has quietly split one
practice into two.

```{r coverage-data}
set.seed(1)
practices <- sprintf("Practice %02d", 1:12)
calls <- do.call(rbind, lapply(seq_along(practices), function(i) {
  scen <- switch(as.character(i %% 3 + 1),
    "1" = c("Straight couple", "Lesbian couple", "Single mother"),
    "2" = c("Straight couple", "Lesbian couple"),
    "3" = "Straight couple")
  data.frame(practice = practices[i], scenario = scen)
}))
head(calls)
```

`mysterycall_scenario_coverage()` reports, per practice, which scenarios were
observed, and rolls that up into complete / partial / singleton counts plus a
missing-per-scenario tally.

```{r coverage}
cov <- mysterycall_scenario_coverage(calls, "practice", "scenario")
cov
head(as.data.frame(cov))
```

A one-character typo in a practice name deflates that count by turning one
practice into two singletons. `mysterycall_flag_near_duplicate_keys()` scans the
keys with an edit distance and surfaces the suspicious pairs for review.

```{r neardup}
keys <- c(practices, "Practice 01 ", "practice 02")  # a trailing space + a case slip
mysterycall_flag_near_duplicate_keys(keys)
```

# 2. A population-average sensitivity model

The subject-specific GLMM (`mysterycall_logistic_model()`) estimates a
*within-practice* effect and needs each practice's random intercept to be
estimable. `mysterycall_gee()` fits the *population-average* companion with a
generalized estimating equation: it uses every call — including the singleton
and dyad practices a matched analysis drops — and reports robust,
correlation-insensitive odds ratios. Agreement between the two is a standard
sensitivity check.

```{r gee, eval = has_geepack}
set.seed(2)
np <- 60L
gee_dat <- data.frame(
  practice = rep(seq_len(np), each = 2),
  scenario = rep(c("Straight couple", "Lesbian couple"), np),
  accepted = rbinom(2 * np, 1, rep(c(0.82, 0.55), np))
)
mysterycall_gee(gee_dat, "accepted", "scenario", "practice",
                reference = c(scenario = "Straight couple"))
```

```{r gee-skip, echo = FALSE, eval = !has_geepack}
cat("geepack not installed; skipping the GEE example.")
```

# 3. Two matched-pair figures

`mysterycall_plot_raincloud()` shows the full outcome distribution by group —
violin, box, and every raw point — the canonical way to present a skewed wait
time without hiding the sample size.

```{r raincloud}
set.seed(3)
wait_dat <- data.frame(
  wait     = c(rpois(60, 10), rpois(60, 24)),
  scenario = rep(c("Commercial", "Medicaid"), each = 60)
)
mysterycall_plot_raincloud(wait_dat, "wait", "scenario",
                           y_lab = "Business days to appointment")
```

`mysterycall_plot_paired_slope()` is the within-practice companion: one line per
practice connecting its calls across scenarios, showing the differencing the
matched design rests on.

```{r slope}
set.seed(4)
pair_dat <- data.frame(
  practice = rep(1:25, each = 2),
  scenario = rep(c("Commercial", "Medicaid"), 25),
  wait     = c(rbind(rpois(25, 9), rpois(25, 20)))
)
mysterycall_plot_paired_slope(pair_dat, "wait", "scenario", "practice",
                              y_lab = "Business days to appointment")
```

# 4. Planning a study: power the count tools do not cover

The package already has power tools for paired, physician-clustered count
outcomes (`mysterycall_marginal_power()`, `mysterycall_twopart_power()`,
`mysterycall_nb_power()`). The ENT access study needed three that fall outside
that: a continuous-outcome calculator, factorial-term power, and covariate-
adjusted power with a higher-level cluster.

## 4a. Continuous outcome under a fixed natural allocation

When only a fraction of practices are rural, equal-allocation formulas
understate the sample size. `mysterycall_ttest_power()` reports both the
equal-allocation N and the total N under the study's real split.

```{r ttest, eval = has_pwr}
mysterycall_ttest_power(mde = c(3, 5, 7), sd = 12, group_frac = 0.15)
```

## 4b. Factorial main effects and interactions

For a saturated `rural × subspecialty × insurance` model,
`mysterycall_lm_interaction_power()` gives the N each term needs at small and
medium effect sizes.

```{r f2, eval = has_pwr}
mysterycall_lm_interaction_power(
  terms = c("rural"                 = 1,
            "subspecialty"          = 6,
            "insurance"             = 1,
            "rural x subspecialty"  = 6,
            "rural x insurance"     = 1),
  n_params = 2 * 7 * 2
)
```

```{r pwr-skip, echo = FALSE, eval = !has_pwr}
cat("pwr not installed; skipping the analytic-power examples.")
```

## 4c. Covariate-adjusted power with a cluster ICC

`mysterycall_adjusted_power()` simulates the *actual adjusted model* — a
negative-binomial GLMM with a rural fixed effect, a subspecialty effect,
adjustment covariates, and a state random intercept whose SD is derived from a
target ICC — instead of guessing a "+20% for clustering" inflation. (A small
`n_sim` is used here so the vignette builds quickly; use several hundred in
practice.)

```{r adjpow, eval = has_glmmTMB}
mysterycall_adjusted_power(
  n_total = 400, rural_frac = 0.15,
  wait_urban = 20, wait_rural = 28, phi = 1.7,
  state_icc = 0.05, n_sim = 20, seed = 1
)
```

```{r adjpow-skip, echo = FALSE, eval = !has_glmmTMB}
cat("glmmTMB not installed; skipping the adjusted-power example.")
```

# 5. Cleaning raw call-log fields

The urogynecology study's raw export had the messy free-text fields every audit
accumulates. These helpers replace the per-study lookup tables that used to
handle them.

`mysterycall_parse_duration()` turns hold-time / call-length free text into a
numeric unit:

```{r duration}
mysterycall_parse_duration(
  c("1.5min", "1min 45 sec", "30sec", "2.75min, 30sec", "O", "3"),
  output_unit = "seconds"
)
```

`mysterycall_clean_zip()` takes the first ZIP when several are present, strips a
ZIP+4 suffix, and restores the leading zero numeric parsing drops:

```{r zip}
mysterycall_clean_zip(c("03110, 03756", 3110, "12345-6789", "no zip", NA))
```

`mysterycall_categorize_wait()` bins a days-to-appointment vector into weekly
categories and the headline ">2-week wait" binary outcome:

```{r catwait}
mysterycall_categorize_wait(c(3, 10, 15, 30, NA), threshold_days = 14)
```

Finally, `mysterycall_link_physicians()` reconciles two physician lists that
share no NPI or other key, via `fastLink` Jaro-Winkler name matching (with
optional blocking on state), returning matched pairs and their posterior match
probability. The example is shown but not run here.

```{r link, eval = FALSE}
roster_a <- data.frame(first_name = c("Katherine", "Robert"),
                       last_name  = c("Smith", "Jones"))
roster_b <- data.frame(first_name = c("Kathryn", "Bob"),
                       last_name  = c("Smith", "Jones"))
mysterycall_link_physicians(roster_a, roster_b,
                            c("first_name", "last_name"), threshold = 0.5)
```

# Summary

| Source study | Functions |
|---|---|
| Matched-pair audit | `mysterycall_scenario_coverage()`, `mysterycall_flag_near_duplicate_keys()`, `mysterycall_gee()`, `mysterycall_plot_raincloud()`, `mysterycall_plot_paired_slope()` |
| ENT access power | `mysterycall_ttest_power()`, `mysterycall_lm_interaction_power()`, `mysterycall_adjusted_power()` |
| Urogynecology data prep | `mysterycall_parse_duration()`, `mysterycall_clean_zip()`, `mysterycall_categorize_wait()`, `mysterycall_link_physicians()` |
