---
title: "Create Drive-Time Isochrones"
author:
  - Tyler Muffly, MD
output:
  rmarkdown::html_vignette:
    df_print: kable
description: >
  Generate drive-time isochrones from geocoded practice coordinates.
vignette: >
  %\VignetteIndexEntry{Create Drive-Time Isochrones}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  chunk_output_type: console
---

> **Moved to `mysterymaps`.** Drive-time isochrone generation and
> isochrone/block-group overlap have moved to the companion
> [`mysterymaps`](https://github.com/mufflyt/mysterymaps) package
> (`mysterymaps::mysterymaps_isochrones_for_df()`,
> `mysterymaps::mysterymaps_calculate_overlap()`). They are no longer part of
> `mysterycall`; this article is kept for reference only.

```{r setup, include=FALSE, purl=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  eval = eval_vignettes
)

if (eval_vignettes) {
  library(mysterycall)
}
```

# Overview

`create_isochrones_for_dataframe()` builds drive-time polygons (isochrones)
from geocoded practice locations using the HERE Routing API. The function is
designed for batch processing: it reads a file of point locations, validates
the coordinates, requests isochrones for each row, and periodically writes
checkpoint files to disk while it runs.

An **isochrone** is a polygon that encloses all points reachable within a given
travel time from a starting location. In a mystery-caller access study, isochrones
answer: *"What is the maximum driving time a patient must accept to reach any
ob-gyn specialist from their home?"*

---

## Before you start

You will need:

1. **A HERE API key** — register at <https://www.here.com/developer> (free tier
   allows ~250,000 isochrone requests/month). Store the key before running:

```{r eval=FALSE, purl=FALSE}
Sys.setenv(HERE_API_KEY = "your-here-key")
```

2. **Geocoded practice locations** — a data frame with `lat` and `long` columns
   in WGS84 decimal degrees. Run `mysterycall_geocode()` first if you are starting
   from street addresses (see `vignette("geocode", package = "mysterycall")`).

3. **A stable output directory** — used for checkpoint files. Do not use
   `tempdir()` on a cluster; choose a persistent path.

---

## Input requirements

The function reads the input file using `easyr::read.any()`, which accepts CSV,
Excel (`.xlsx`/`.xls`), and several other tabular formats. Required columns:

| Column | Type | Description |
|--------|------|-------------|
| `lat`  | numeric | Latitude in decimal degrees (WGS84) |
| `long` | numeric | Longitude in decimal degrees (WGS84) |

Rows where either coordinate is `NA` or falls outside US bounds
(lat 18°–72°, long −180° to −65°) are rejected before any API call is made.
Include an NPI or another unique identifier column so polygons can be traced
back to the originating provider.

---

## Minimal workflow

The typical four-step pattern:

```{r eval=FALSE, purl=FALSE}
library(mysterycall)

# 1. Geocode practice addresses → lat/long
geocoded <- mysterycall_geocode(
  file_path           = "data/providers_with_addresses.csv",
  google_maps_api_key = Sys.getenv("GOOGLE_MAPS_API_KEY"),
  output_file_path    = "data/geocoded_providers.csv"
)

# 2. Rename coordinate columns to lat / long
iso_input <- geocoded |>
  dplyr::rename(lat = latitude, long = longitude)

# 3. Write to a file (easyr::read.any() will read it back)
readr::write_csv(iso_input, "data/iso_input.csv")

# 4. Generate isochrones
isochrones_sf <- create_isochrones_for_dataframe(
  input_file    = "data/iso_input.csv",
  breaks        = c(30 * 60, 60 * 60, 120 * 60, 180 * 60),  # 30/60/120/180 min
  api_key       = Sys.getenv("HERE_API_KEY"),
  output_dir    = "data/isochrone_checkpoints",
  save_interval = 100   # save checkpoint every 100 providers
)
```

---

## Understanding the key arguments

| Argument | Default | Description |
|----------|---------|-------------|
| `input_file` | — | Path to the CSV / Excel file with `lat` and `long` columns |
| `breaks` | — | Travel-time cut points **in seconds** (e.g. `1800` = 30 min) |
| `api_key` | `HERE_API_KEY` env var | HERE API key for the Routing API |
| `output_dir` | — | Directory where checkpoint `.rds` and `.gpkg` files are written |
| `save_interval` | 240 | Save a checkpoint after every N successfully processed rows |

**Choosing `breaks`:** Common study designs use 30, 60, 120, and 180 minutes.
The HERE API generates one polygon per cut point per location, so four breaks
on 500 providers = 2,000 API calls. The free tier accommodates this with margin.

---

## What gets returned

The function returns a combined `sf` object in CRS 4326 (WGS84). Each row is
one isochrone polygon. Key added columns:

| Column | Description |
|--------|-------------|
| `point_index` | Row number of the source location in the input file |
| `travel_time_minutes` | The break value expressed in minutes |
| `name` | Provider name or identifier carried over from the input |
| `column_label` | Human-readable label, e.g. `"30 min"` |
| all input columns | All non-geometry columns from the source row are retained |

```{r eval=FALSE, purl=FALSE}
# Inspect the result
names(isochrones_sf)

dplyr::count(isochrones_sf, travel_time_minutes)
#>   travel_time_minutes   n
#> 1                  30 498
#> 2                  60 498
#> 3                 120 497
#> 4                 180 497
# (some providers may return fewer polygons if the API cannot reach them)

# Save the full sf object
sf::st_write(isochrones_sf, "data/isochrones.gpkg", delete_dsn = TRUE)
readr::write_rds(isochrones_sf, "data/isochrones.rds")
```

---

## Checkpoint files and resuming interrupted runs

Long batches (500+ providers) can take an hour or more. The function writes
two checkpoint files to `output_dir` after every `save_interval` rows:

- `isochrones_checkpoint_<timestamp>.rds` — the combined sf object so far
- `isochrones_checkpoint_<timestamp>.gpkg` — same data in GeoPackage format

If a run is interrupted (network error, cluster timeout), load the most recent
checkpoint and pass the completed `point_index` values so they are skipped:

```{r eval=FALSE, purl=FALSE}
# Load the most recent checkpoint
checkpoints <- list.files("data/isochrone_checkpoints", pattern = "\\.rds$",
                           full.names = TRUE)
partial <- readr::read_rds(sort(checkpoints, decreasing = TRUE)[1])

completed_idx <- unique(partial$point_index)
message("Resuming from row ", max(completed_idx) + 1,
        " (", length(completed_idx), " rows already done)")

# Pass the source file again; the function will skip completed indices
isochrones_sf <- create_isochrones_for_dataframe(
  input_file    = "data/iso_input.csv",
  breaks        = c(30 * 60, 60 * 60, 120 * 60, 180 * 60),
  api_key       = Sys.getenv("HERE_API_KEY"),
  output_dir    = "data/isochrone_checkpoints",
  save_interval = 100,
  skip_indices  = completed_idx   # omit already-processed rows
)

# Combine the checkpoint and the new results
isochrones_sf <- dplyr::bind_rows(partial, isochrones_sf)
```

---

## Practical guidance

- **Test with 5–10 rows first** to confirm the API key, coordinate format, and
  output directory are correct before submitting the full batch.
- **Use street-level geocoding.** ZIP-code centroid coordinates produce
  isochrones centered on the post-office, not the clinic — a meaningful spatial
  error in dense urban areas.
- **Choose a persistent `output_dir`**, not `tempdir()`, so checkpoints survive
  a session restart or cluster requeue.
- **Rate limits:** HERE's free tier allows 5 requests/second. The function
  includes a small delay between requests; do not remove it.
- **CRS:** all output is in EPSG 4326. Transform to a projected CRS
  (e.g. EPSG 5070 Albers Equal Area) before computing areas or intersections.

---

# Next step: population overlap analysis

After generating isochrones, overlay them on Census block group geometries to
calculate the reachable female population within each travel-time band:

```{r eval=FALSE, purl=FALSE}
# Load the enriched block group layer (built in the Census vignette)
bg_sf <- readr::read_rds("data/block_groups_with_demographics.rds")

# Calculate intersection overlap and save results
overlap_results <- calculate_intersection_overlap_and_save(
  isochrones_sf = isochrones_sf,
  bg_sf         = bg_sf,
  output_dir    = "data/overlap_results"
)
```

See `vignette("get_census_data", package = "mysterycall")` for building `bg_sf`,
and `vignette("mapping", package = "mysterycall")` for visualizing the output.
