---
title: "Data Quality: Phone Validation, Name Parsing, and Safe Joins"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    df_print: kable
description: >
  How to validate NANP phone numbers, parse physician names into structured
  components, and perform audited joins that guard against row duplication
  and coverage loss.
vignette: >
  %\VignetteIndexEntry{Data Quality: Phone Validation, Name Parsing, and Safe Joins}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = eval_vignettes
)
library(mysterycall)
```

## Overview

Mystery-caller studies assemble provider rosters from several upstream sources
(NPPES, CMS Care Compare, internal call logs) that each carry their own
formatting quirks. Three categories of data-quality problems arise repeatedly:

1. **Phone numbers** — formatting noise (`(303) 555-1234`, `303.555.1234`,
   `+13035551234`) and area codes that don't match the provider's reported
   state.
2. **Physician names** — a single free-text name field that must be split into
   first / middle / last / suffix components before being used in call scripts
   or de-duplication logic.
3. **Table joins** — silent row multiplication when the right-hand table is not
   unique on the join key, or silent data loss when a large fraction of left-hand
   rows find no match on the right.

This vignette walks through the `mysterycall` helpers that address each problem.
All code runs without any API keys using only built-in package data.

---

## 1. Phone validation (`mysterycall_validate_phone`)

### Background

The North American Numbering Plan (NANP) defines the structure of US phone
numbers:

```
NPA-NXX-XXXX
```

where **NPA** (Numbering Plan Area, the "area code") must have a first digit of
2–9, and **NXX** (the central-office code) must also have a first digit of 2–9
and must not be an N11 service code (211, 311, …, 911).

`mysterycall_validate_phone()` strips formatting, enforces those structural
rules, and — optionally — checks that the area code belongs to the provider's
reported practice state using the bundled `inst/extdata/nanp_area_codes_us.csv`
lookup table.

### Basic usage

```{r phone-basic}
# A well-formed Colorado number
mysterycall_validate_phone("(303) 555-1234")

# Missing number
mysterycall_validate_phone(NA_character_)

# Non-NANP format (too few digits)
mysterycall_validate_phone("555-1234")
```

The function always returns a data frame with five columns:

| Column | Type | Meaning |
|--------|------|---------|
| `phone_e164_valid` | logical | `TRUE` if the number passes all NANP rules |
| `phone_npa` | character | The three-digit area code, or `NA` |
| `phone_state_from_npa` | character | State associated with the NPA in the lookup table |
| `phone_area_code_matches_state` | logical | `TRUE` if NPA ↔ `practice_state` agree; `NA` if no state was supplied |
| `phone_validity_flag` | character | One of the five validity codes below |

| Flag | Meaning |
|------|---------|
| `valid` | Passes all NANP rules; area code matches practice state (if supplied) |
| `missing` | Input is `NA` or empty |
| `invalid_format` | Cannot be reduced to a 10-digit NANP number |
| `unknown_area_code` | Digits are structurally valid but NPA is not in the lookup table |
| `area_code_state_mismatch` | NPA is valid but registered to a different state |

### State-mismatch detection

Pass the provider's two-letter state abbreviation to catch numbers whose area
code is registered to a different state:

```{r phone-state}
# Colorado provider with a New York area code
mysterycall_validate_phone("(212) 555-0199", practice_state = "CO")
```

### Validating a full roster column

```{r phone-roster}
roster <- data.frame(
  npi            = c("1234567890", "0987654321", "1111111111"),
  phone          = c("(720) 555-0101", "999-0000", "(303) 867-5309"),
  practice_state = c("CO",            "CO",        "CO"),
  stringsAsFactors = FALSE
)

# vectorised: one call per row, returns a data frame
phone_results <- do.call(rbind, Map(
  mysterycall_validate_phone,
  phone_str      = roster$phone,
  practice_state = roster$practice_state
))

cbind(roster["npi"], phone_results[, c("phone_e164_valid", "phone_validity_flag")])
```

The NANP lookup table is lazy-loaded once per session and cached internally,
so repeated calls in a loop incur no extra I/O.

---

## 2. Name parsing (`mysterycall_parse_physician_name`)

### Background

Provider names arrive from NPPES and call logs in several formats:

- `"Jane A. Smith, MD"` — credential after name
- `"Smith, Jane A."` — last-first comma format (two-part)
- `"Smith, Jane, Jr."` — last-first-suffix three-part comma format
- `"María de la Cruz"` — name particle in the last name
- `"Robert Smith DO"` — `DO` credential that must not become `last_name`
- `"Linda Do"` — Vietnamese surname `Do` that must **not** be treated as a credential

`mysterycall_parse_physician_name()` wraps `humaniformat::parse_names()` and
applies deterministic post-processing rules for these edge cases, then assigns
a confidence level (`high` / `medium` / `low`) so downstream consumers can
flag low-confidence parses for manual review.

### Single name

```{r parse-single}
mysterycall_parse_physician_name("Smith, Jane, Jr.")
```

### Vectorised over a column

```{r parse-vector}
raw_names <- c(
  "Maria de la Cruz",
  "Smith, John, Jr.",
  "Robert Smith DO",
  "Linda Do",
  NA_character_
)

parsed <- mysterycall_parse_physician_name(raw_names)
parsed[, c("first_name", "last_name", "suffix", "parse_confidence", "parse_warnings")]
```

### Using the built-in `physicians` dataset

```{r parse-physicians}
# Parse the 4,659 physician names from the bundled roster
parsed_physicians <- mysterycall_parse_physician_name(
  mysterycall::physicians$name
)

# Confidence breakdown
table(parsed_physicians$parse_confidence)
```

### Validating parsed output

After parsing a full roster, call `mysterycall_validate_parsed_names()` to
flag structural quality problems in each row:

```{r parse-validate}
validated <- mysterycall_validate_parsed_names(parsed)

# Overview of quality issues
table(validated$quality_issue, useNA = "ifany")

# Rows that are not valid (missing first or last name)
validated[!validated$is_valid, c("first_name", "last_name",
                                  "parse_confidence", "quality_issue")]
```

Rows with `parse_confidence == "low"` or a non-`NA` `quality_issue` should be
reviewed manually before use in call scripts.

### Formatting names for output

Once parsed, use `mysterycall_format_physician_name()` to render names in a
consistent house style:

```{r parse-format}
mysterycall_format_physician_name(
  first_name     = "Jane",
  last_name      = "Smith",
  middle_name    = "A.",
  suffix         = "MD",
  format         = "last_first",
  include_suffix = TRUE
)

# Vectorised
mysterycall_format_physician_name(
  first_name  = c("John",  "Maria"),
  last_name   = c("Smith", "Cruz"),
  suffix      = c("MD",    NA),
  format      = "first_last"
)
```

### Built-in edge-case test suite

The function ships with 13 known-difficult cases (DO credential, Vietnamese
surnames, three-part comma, hyphenated names) that you can run at any time to
verify parser behavior after an upgrade:

```{r parse-test}
tbl <- mysterycall_test_name_parser()
cat(sprintf("%d / %d high-confidence valid parses\n",
            sum(tbl$is_valid & tbl$parse_confidence == "high"),
            nrow(tbl)))
```

---

## 3. Safe joins (`mysterycall_safe_left_join` and friends)

### The problem with silent join failures

A plain `dplyr::left_join()` will silently:

- **Multiply rows** if the right-hand key is not unique (a many-to-one match
  becomes many-to-many).
- **Drop rows** if the join key types don't match (e.g., integer vs. character
  NPI), causing every row to land in the unmatched group.
- **Return no warning** when 40 % of left-hand NPIs have no match on the right.

The `mysterycall_safe_*_join()` family wraps `dplyr` joins with pre- and
post-join assertions that make these failures loud.

### `mysterycall_safe_left_join`

```{r join-left}
calls <- data.frame(
  npi  = c("1111111111", "2222222222", "3333333333"),
  date = as.Date(c("2025-01-10", "2025-01-11", "2025-01-12")),
  stringsAsFactors = FALSE
)

demographics <- data.frame(
  npi    = c("1111111111", "2222222222"),
  gender = c("Female", "Male"),
  stringsAsFactors = FALSE
)

# 2 of 3 left rows match → coverage = 66.7 % < 95 % threshold → error
tryCatch(
  suppressMessages(mysterycall_safe_left_join(
    left         = calls,
    right        = demographics,
    by           = "npi",
    label_left   = "call_log",
    label_right  = "demographics",
    min_coverage = 0.95
  )),
  error = function(e) message("Caught: ", conditionMessage(e))
)

# Lower threshold to match the actual coverage
result <- suppressMessages(mysterycall_safe_left_join(
  left         = calls,
  right        = demographics,
  by           = "npi",
  label_left   = "call_log",
  label_right  = "demographics",
  min_coverage = 0.60
))
result
```

### Key-type harmonization

If one table stores NPI as integer and the other as character, a plain join
silently returns zero matches. The safe join coerces both sides to character:

```{r join-types}
left2  <- data.frame(npi = 1111111111L, stringsAsFactors = FALSE)
right2 <- data.frame(npi = "1111111111", x = 42L, stringsAsFactors = FALSE)

suppressMessages(mysterycall_safe_left_join(
  left2, right2, by = "npi",
  label_left  = "left",
  label_right = "right",
  min_coverage = 0
))
```

### Asserting uniqueness before a join

`mysterycall_assert_unique_keys()` lets you verify that a table is unique on
one or more columns before passing it to any join:

```{r join-assert}
dupes <- data.frame(
  npi    = c("1111111111", "1111111111"),
  source = c("NPPES", "CMS"),
  stringsAsFactors = FALSE
)

# Error: found 1 duplicate group
tryCatch(
  mysterycall_assert_unique_keys(dupes, key_cols = "npi", label = "demographics"),
  error = function(e) message("Caught: ", conditionMessage(e))
)

# Pass dedupe = TRUE to drop duplicates instead:
clean <- suppressMessages(
  mysterycall_assert_unique_keys(dupes, key_cols = "npi",
                                  label = "demographics", dedupe = TRUE)
)
nrow(clean)
```

### Inner, semi, and anti joins

```{r join-variants}
# Inner join — only matched rows; defaults to min_coverage = 0.90
result_inner <- suppressMessages(mysterycall_safe_inner_join(
  left        = calls,
  right       = demographics,
  by          = "npi",
  label_left  = "call_log",
  label_right = "demographics",
  min_coverage = 0.60
))
nrow(result_inner)   # 2 rows: only matched NPIs

# Semi join — keep left rows that match; no right columns added
matched_calls <- suppressMessages(mysterycall_safe_semi_join(
  left        = calls,
  right       = demographics,
  by          = "npi",
  label_left  = "call_log",
  label_right = "demographics",
  min_coverage = 0.60
))
matched_calls   # 2 rows, same columns as calls

# Anti join — keep left rows with NO match
unmatched_calls <- suppressMessages(mysterycall_safe_anti_join(
  left        = calls,
  right       = demographics,
  by          = "npi",
  label_left  = "call_log",
  label_right = "demographics"
))
unmatched_calls   # 1 row: NPI 3333333333
```

---

## 4. Putting it together: a complete roster QA pipeline

The built-in `physicians` dataset contains 4,659 OBGYN subspecialists with NPI
and name. The block below demonstrates the full data-quality pipeline without
any API calls.

```{r pipeline}
library(dplyr)

# ── 1. Parse physician names ──────────────────────────────────────────────────
roster <- mysterycall::physicians |>
  select(NPI, name, subspecialty)

parsed_names <- mysterycall_parse_physician_name(roster$name)
roster <- bind_cols(roster, parsed_names)

low_conf <- filter(roster, parse_confidence == "low")
message(sprintf("%d name(s) flagged as low-confidence (%.1f%%)",
                nrow(low_conf), 100 * nrow(low_conf) / nrow(roster)))

# ── 2. Validate parsed names ──────────────────────────────────────────────────
validated_names <- mysterycall_validate_parsed_names(parsed_names)
issues <- filter(validated_names, !is.na(quality_issue))
message(sprintf("%d name quality issue(s) detected:", nrow(issues)))
print(table(issues$quality_issue))

# ── 3. Illustrate a safe join with a synthetic demographics sidecar ───────────
# In a real study this would be CMS Care Compare or similar
demographics_sidecar <- data.frame(
  NPI    = roster$NPI[1:4000],       # 4,000 of 4,659 have demographics
  gender = sample(c("Female", "Male", "Unknown"),
                  4000, replace = TRUE, prob = c(.7, .28, .02)),
  stringsAsFactors = FALSE
)

roster_enriched <- suppressMessages(
  mysterycall_safe_left_join(
    left         = roster,
    right        = demographics_sidecar,
    by           = "NPI",
    label_left   = "physicians",
    label_right  = "demographics_sidecar",
    min_coverage = 0.85   # expect ≥ 85 % match rate
  )
)

# Coverage summary
cat(sprintf(
  "Roster rows: %d  |  Enriched rows: %d  |  Coverage: %.1f%%\n",
  nrow(roster), nrow(roster_enriched),
  100 * mean(!is.na(roster_enriched$gender))
))
```
