---
title: "Data Cleaning: A Complete Workflow for Mystery Caller Studies"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    df_print: kable
description: >
  End-to-end data cleaning for mystery caller studies: NPI validation, phone
  validation, address normalization, Phase 1 and Phase 2 log cleaning,
  duplicate detection, and clinician data retrieval.
vignette: >
  %\VignetteIndexEntry{Data Cleaning: A Complete Workflow for Mystery Caller Studies}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE, purl=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = eval_vignettes
)
library(mysterycall)
```

## Overview

A mystery caller study typically produces two raw datasets:

1. **Phase 1** — the call scheduling log (physician name, practice, phone, state, NPI, insurance type)
2. **Phase 2** — the call outcome log (wait time, appointment date, call result)

Both datasets need cleaning before analysis. This vignette walks through every
`mysterycall` cleaning tool in the order you would typically apply them, using
only synthetic data so no API keys are required.

---

## 1. NPI validation (`mysterycall_validate_npi`)

National Provider Identifier (NPI) numbers are the primary join key across all
downstream tables. Bad NPIs must be caught early — a row with a wrong NPI will
silently match the wrong physician in every subsequent lookup.

`mysterycall_validate_npi()` applies three checks in order:

1. Strip non-digit characters and leading/trailing whitespace.
2. Require exactly 10 digits.
3. Validate the Luhn checksum (the algorithm the NPI standard mandates).

```{r npi-validation, purl=FALSE}
raw <- data.frame(
  npi           = c("1234567893", "0000000000", "123456789X", NA, "1234567893 "),
  physician_name = c("Smith, Alice", "Jones, Bob", "Park, Carol",
                     "Lee, David", "Smith, Alice"),
  stringsAsFactors = FALSE
)

validated <- mysterycall_validate_npi(raw)
validated[, c("npi", "npi_is_valid", "physician_name")]
```

Only rows where `npi_is_valid == TRUE` survive. The returned data frame is a
strict subset of the input — no rows are ever added.

---

## 2. Phone validation (`mysterycall_validate_phone`)

Office phone numbers collected during calls come in many formats. Before you
can use them for scheduling or re-call logic, validate each number against
North American Numbering Plan (NANP) rules and optionally cross-check the
area code against the physician's reported state.

```{r phone-validation, purl=FALSE}
phones <- c(
  "(303) 555-1212",   # valid CO number
  "800-555-0199",     # toll-free — valid syntax, no state match
  "555-1212",         # missing area code — invalid
  NA,
  "13035551212"       # with leading country code — should strip to 10 digits
)

phone_results <- mysterycall_validate_phone(
  phones,
  practice_state = "CO"
)
phone_results
```

The `phone_validity_flag` column distinguishes five outcomes:

| Flag | Meaning |
|------|---------|
| `"valid"` | Passes all NANP rules and area code matches state |
| `"missing"` | NA or blank |
| `"invalid_format"` | Wrong digit count or invalid NPA/NXX digits |
| `"unknown_area_code"` | Syntactically valid but area code not in lookup table |
| `"area_code_state_mismatch"` | Valid syntax, known area code, but wrong state |

Pass `practice_state = NULL` to skip the state-matching check when you only
want syntax validation.

---

## 3. Address normalization

Addresses from NPPES and practice websites use inconsistent abbreviations
("Street" vs "St", "Suite" vs "Ste", "California" vs "CA"). The normalizer
functions standardize each component to USPS format so addresses can be
matched or geocoded reliably.

Apply the functions in this order:

### 3a. ASCII normalization

```{r addr-ascii, eval=FALSE, purl=FALSE}
# mysterycall_ascii_norm() is an internal helper; shown here for illustration.
raw_addr <- "123 Cafe Boulevard, Suite 4A"
mysterycall:::mysterycall_ascii_norm(raw_addr)
```

### 3b. Directional normalization (North → N, Northeast → NE)

```{r addr-directionals, eval=FALSE, purl=FALSE}
mysterycall_normalize_directionals("1400 North Main Street")
mysterycall_normalize_directionals("201 Southeast Oak Avenue")
```

### 3c. Street suffix normalization (Street → ST, Boulevard → BLVD)

```{r addr-suffix, eval=FALSE, purl=FALSE}
mysterycall_normalize_suffix("100 MAIN STREET")
mysterycall_normalize_suffix("500 OAK BOULEVARD SUITE 3")
```

### 3d. Unit designator normalization (Suite → STE, Apartment → APT)

```{r addr-units, eval=FALSE, purl=FALSE}
mysterycall_normalize_units("100 Main St Suite 4", addr2 = NA)
mysterycall_normalize_units("200 Elm Ave", addr2 = "Apartment 7B")
```

### 3e. State abbreviation normalization

```{r addr-state, eval=FALSE, purl=FALSE}
mysterycall_normalize_state("California")
mysterycall_normalize_state("new york")
mysterycall_normalize_state("CO")   # already abbreviated — returned as-is
```

### 3f. ZIP code normalization (zero-pad to 5 digits)

```{r addr-zip, eval=FALSE, purl=FALSE}
mysterycall_normalize_zip5(c("80202", "1234", "12345-6789", NA))
```

### 3g. P.O. Box detection

Flag addresses that cannot be visited in person:

```{r addr-po-box, eval=FALSE, purl=FALSE}
mysterycall_is_po_box(c("P.O. Box 123", "100 Main Street", "PO Box 99", NA))
```

### Full address pipeline

Pipe all steps together for a data frame:

```{r addr-pipeline, eval=FALSE, purl=FALSE}
df_addr <- data.frame(
  addr1 = c("1234 north main street", "500 Oak Blvd Suite 3A",
             "P.O. Box 77", "789 Southeast Elm Avenue"),
  addr2 = c(NA, NA, NA, "Room 12"),
  state = c("Colorado", "california", "NY", "Texas"),
  zip   = c("80202", "90210", "10001", "77001"),
  stringsAsFactors = FALSE
)

df_addr$addr1_norm  <- sapply(df_addr$addr1, function(a) {
  a |>
    mysterycall_ascii_norm() |>
    mysterycall_normalize_directionals() |>
    mysterycall_normalize_suffix() |>
    mysterycall_strip_suite()
})
df_addr$state_abbr  <- mysterycall_normalize_state(df_addr$state)
df_addr$zip5        <- mysterycall_normalize_zip5(df_addr$zip)
df_addr$is_po_box   <- mysterycall_is_po_box(df_addr$addr1)

df_addr[, c("addr1_norm", "state_abbr", "zip5", "is_po_box")]
```

---

## 4. Phase 1 log cleaning (`mysterycall_clean_phase1`)

The Phase 1 log is the scheduling dataset — one row per physician per
insurance type called. `mysterycall_clean_phase1()` performs:

- Column name standardization via `janitor::clean_names()`
- Text squishing and phone formatting
- NPI preservation and random-ID generation for rows without NPIs
- Optional row duplication for insurance pairing
- Academic/private practice classification
- Audit trail generation (JSON) for data provenance

```{r clean-phase1, purl=FALSE}
phase1_raw <- data.frame(
  Names         = c("Smith, Alice MD", "Jones, Bob DO", "Park, Carol MD"),
  practice_name = c("Denver University Medical", "Downtown Family Clinic",
                    "Rocky Mountain Surgery"),
  phone_number  = c("3035551212", "(720) 555-9999", "303-555-8888"),
  state_name    = c("Colorado", "Colorado", "Colorado"),
  npi           = c("1234567893", NA, "1578798779"),
  for_redcap    = c(NA, NA, NA),
  stringsAsFactors = FALSE
)

result_phase1 <- mysterycall_clean_phase1(
  phase1_raw,
  output_directory = tempdir(),
  verbose          = FALSE,
  duplicate_rows   = TRUE
)

result_phase1[, c("id", "dr_name", "insurance", "phone_number", "academic",
                  "processing_flag_generated_id")]
```

The function invisibly returns the cleaned data frame and writes two files to
`output_directory`:

- `clean_phase_1_results_<timestamp>.csv` (or `.parquet` if `output_format = "parquet"`)
- `audit_trail_<timestamp>.json` — machine-readable provenance record

### Accessing the audit trail

```{r phase1-audit, purl=FALSE}
trail <- attr(result_phase1, "audit_trail")
cat(sprintf(
  "Rows: %d → %d  |  Duration: %.2fs  |  NPI completeness: %.0f%%\n",
  trail$input_rows,
  trail$output_rows,
  trail$duration_seconds,
  trail$quality_metrics$completeness_npi * 100
))
```

---

## 5. Duplicate detection (`mysterycall_check_duplicates`)

Physicians called more than the protocol allows inflate your sample and bias
results. Flag repeat calls before analysis:

```{r duplicates, purl=FALSE}
calls <- data.frame(
  npi         = c("1234567893", "1234567893", "1234567893",
                  "9876543210", "9876543210"),
  call_date   = as.Date(c("2024-01-10", "2024-01-12", "2024-01-15",
                           "2024-01-10", "2024-01-11")),
  insurance   = c("BCBS", "Medicaid", "BCBS",
                  "BCBS", "Medicaid"),
  stringsAsFactors = FALSE
)

dup_report <- mysterycall_check_duplicates(calls, id_col = "npi", max_calls = 2L)
dup_report
```

Rows flagged by `mysterycall_check_duplicates()` should be reviewed and
removed (or justified) before the primary analysis.

---

## 6. Column renaming (`mysterycall_rename_columns`)

Raw exports from REDCap or spreadsheets often have verbose or inconsistent
column names. Use `mysterycall_rename_columns()` to rename by substring match:

```{r rename, purl=FALSE}
messy <- data.frame(
  physician_full_name_as_entered = "Smith, Alice",
  office_telephone_number        = "303-555-1212",
  national_provider_identifier   = "1234567893",
  stringsAsFactors = FALSE
)

clean <- mysterycall_rename_columns(
  messy,
  target_strings = c("physician_full_name", "telephone", "national_provider"),
  new_names      = c("physician_name", "phone", "npi")
)
names(clean)
```

---

## 7. Clinician data retrieval (`mysterycall_get_clinician_data`)

Once NPIs are validated, enrich your roster with NPPES data (specialty,
address, credential) via the `provider` package. This step requires the
`provider` package to be installed from GitHub:

```r
remotes::install_github("andrewallenbruce/provider")
```

```{r clinician-data, purl=FALSE}
npi_df <- data.frame(
  npi = c("1234567893", "1578798779"),
  stringsAsFactors = FALSE
)

# Returns NULL silently per NPI when 'provider' is not installed.
clinician_info <- mysterycall_get_clinician_data(npi_df)
clinician_info
```

The returned tibble has one row per NPI and carries columns such as
`taxonomies_desc` (NPPES taxonomy code), `basic_first_name`, `basic_last_name`,
and practice address fields.

> **Important:** Do not use `taxonomies_desc` to assign subspecialty. NPPES
> taxonomy codes reflect broad specialty groupings only. Use
> `mysterycall_parse_certification_subspecialty()` for subspecialty
> assignment.

---

## Putting it all together

A complete cleaning pipeline for a new mystery caller dataset:

```{r full-pipeline, eval = FALSE, purl=FALSE}
library(mysterycall)

# 1. Validate NPIs
raw      <- readr::read_csv("phase1_raw.csv")
valid    <- mysterycall_validate_npi(raw)

# 2. Clean Phase 1 log
phase1   <- mysterycall_clean_phase1(
  valid,
  output_directory = "output/",
  verbose          = TRUE
)

# 3. Validate phone numbers
phone_ok <- mysterycall_validate_phone(
  phase1$phone_number,
  practice_state = phase1$state_name
)
phase1   <- cbind(phase1, phone_ok)

# 4. Normalize addresses
phase1$addr1_clean <- with(phase1, {
  mysterycall_normalize_directionals(
    mysterycall_normalize_suffix(
      mysterycall_ascii_norm(address_line_1)
    )
  )
})
phase1$state_abbr <- mysterycall_normalize_state(phase1$state_name)

# 5. Detect duplicates
dups     <- mysterycall_check_duplicates(phase1, id_col = "npi", max_calls = 2L)
phase1   <- phase1[phase1$npi %in% dups$npi[dups$n_calls <= 2L], ]

# 6. Enrich with NPPES clinician data (requires 'provider' package)
enriched <- mysterycall_get_clinician_data(phase1)
```

---

## Function reference

| Function | Purpose |
|----------|---------|
| `mysterycall_validate_npi()` | Luhn-validate NPI numbers; drop invalid rows |
| `mysterycall_validate_phone()` | NANP syntax check + area-code-to-state match |
| `mysterycall_ascii_norm()` | Strip non-ASCII characters and collapse whitespace |
| `mysterycall_normalize_directionals()` | Expand/abbreviate N/S/E/W/NE/… |
| `mysterycall_normalize_suffix()` | Street → ST, Boulevard → BLVD, etc. |
| `mysterycall_normalize_units()` | Suite → STE, Apartment → APT, etc. |
| `mysterycall_normalize_state()` | Full state name → 2-letter abbreviation |
| `mysterycall_normalize_zip5()` | Zero-pad ZIP to 5 digits |
| `mysterycall_strip_suite()` | Remove suite/unit token from address line 1 |
| `mysterycall_is_po_box()` | Detect P.O. Box addresses |
| `mysterycall_has_street_number()` | Detect addresses that include a building number |
| `mysterycall_clean_phase1()` | Full Phase 1 log cleaning with audit trail |
| `mysterycall_check_duplicates()` | Flag physicians called more than N times |
| `mysterycall_rename_columns()` | Rename columns by substring match |
| `mysterycall_get_clinician_data()` | Retrieve NPPES data via `provider` package |
