---
title: "Guarding Against a Contaminated Outcome Variable"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    df_print: kable
description: >
  How a wait-time column becomes fill-down contaminated, why the damage is
  invisible in summary statistics, and how to refuse the column outright
  instead of modelling it.
vignette: >
  %\VignetteIndexEntry{Guarding Against a Contaminated Outcome Variable}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE, purl=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = eval_vignettes
)
library(mysterycall)
```

## The defect this vignette is about

A mystery-caller log has one row per call. Some calls reach an office and get an
appointment date; most of the rest never reach anyone at all. The wait-time
column should therefore be missing for every call that produced no appointment.

Fill-down contamination is what happens when that column gets carried forward.
It usually arrives through a spreadsheet: the rows are sorted so that each
answered call is followed by the calls that were not, someone drags a formula
down or applies a fill operation, and every unreachable office silently inherits
the wait of the last office that answered.

The reason this deserves its own vignette, rather than a line in a checklist, is
that **the result does not look broken**.

```{r contaminated-frame}
calls <- data.frame(
  clinic = c("Northside", "Lakeview", "Harbor",
             "Cedar Park", "Ridgeway",
             "Summit", "Fairmont", "Brookline", "Eastgate"),
  appointment_date = as.Date(c("2019-07-11", NA, NA,
                               "2019-08-23", NA,
                               "2019-09-03", NA, NA, NA)),
  reason_for_exclusions = c("Able to contact", "Went to voicemail", "Wrong number listed",
                            "Able to contact", "Went to voicemail",
                            "Able to contact", "Wrong number listed",
                            "Went to voicemail", "Went to voicemail"),
  business_days_until_appointment = c(8, 8, 8, 16, 16, 28, 28, 28, 28)
)
calls
```

Every value is a plausible number of business days. Nothing is negative, nothing
is absurdly large, and the column is complete, which usually reads as a sign of
careful data collection rather than a warning.

## Why summary statistics will not save you

The natural instinct is to look at the distribution. That is exactly what fails
here, because the fabricated values are copies of real ones and so share their
distribution.

```{r means}
honest <- calls$business_days_until_appointment[!is.na(calls$appointment_date)]
c(contaminated_mean = mean(calls$business_days_until_appointment),
  honest_mean       = mean(honest))
```

The two means are close, and in the real dataset that motivated this function
they were closer still: 23.8 against an honest 23.0. A reviewer comparing that
number to the published literature would find nothing to object to. The mean is
not wrong because the arithmetic is wrong; it is wrong because it is computed
over rows that never had an appointment.

The damage only becomes visible when you stop asking what the numbers are and
start asking **which rows they are attached to**.

```{r pairing}
data.frame(
  clinic       = calls$clinic,
  wait         = calls$business_days_until_appointment,
  has_appt     = !is.na(calls$appointment_date),
  was_excluded = calls$reason_for_exclusions != "Able to contact"
)
```

Six of the nine rows carry a wait time despite having no appointment date, and
four of them were never contacted at all. An office that went to voicemail
cannot have been given an appointment in eleven business days.

## Refusing the column

`mysterycall_guard_contaminated_wait()` stops rather than reports. That is the
difference between it and `mysterycall_flag_exclusion_discrepancy()`, which
returns the offending rows so a human can adjudicate them. A guard is for the
case where you have already decided the column is not analysable and want the
pipeline to fail if anyone forgets.

```{r guard-fails, error = TRUE}
mysterycall_guard_contaminated_wait(
  calls,
  appointment_col = "appointment_date",
  exclusion_col   = "reason_for_exclusions",
  contact_value   = "Able to contact"
)
```

It checks three signatures, in decreasing order of certainty:

1. **A wait with no appointment date.** An office that never gave a date cannot
   have a measured wait. This is the strongest evidence and needs no assumptions.
2. **A wait on an excluded call.** The call is recorded as not having produced an
   appointment, so a wait attached to it contradicts the exclusion.
3. **A carry-forward run.** Consecutive identical waits where the later row has
   no appointment of its own to justify them. This is the fingerprint of the
   fill operation itself.

The condition carries the offending rows, so you can triage rather than guess:

```{r guard-condition}
cond <- tryCatch(
  mysterycall_guard_contaminated_wait(calls, appointment_col = "appointment_date"),
  mysterycall_contaminated_wait = function(e) e
)
cond$rows
cond$counts
```

## A guard that cannot fail is worse than no guard

Two behaviours are worth knowing about, because both exist to stop the function
from giving false assurance.

**Repeated values are not automatically suspicious.** Two offices can genuinely
offer the same wait. The guard only objects when the repeat is *unjustified*,
meaning the later row has no appointment of its own.

```{r legitimate-repeats}
legit <- data.frame(
  appointment_date                = as.Date(c("2019-07-11", "2019-07-12", "2019-07-15")),
  business_days_until_appointment = c(10, 10, 10)
)
mysterycall_guard_contaminated_wait(legit, appointment_col = "appointment_date")
```

**A non-numeric wait column is refused, not passed.** Call logs of this kind
often carry a banded version of the wait alongside the numeric one. Returning a
clean bill of health for a column the guard never examined would be exactly the
false assurance it exists to prevent, so it errors instead.

```{r banded, error = TRUE}
banded <- data.frame(
  appointment_date                = as.Date(c("2019-07-11", NA)),
  business_days_until_appointment = c("1 to 10 business days", NA)
)
mysterycall_guard_contaminated_wait(banded, appointment_col = "appointment_date")
```

## What to do instead

Do not repair the column, and do not delete it.

Repairing it means guessing which values were real, and the fill operation has
already destroyed the information needed to tell. Deleting it means the next
person to open the source file finds an undocumented column and no reason not to
use it, which is how a defect gets reintroduced three years later.

Derive the outcome from the source fields instead. The call date and the
appointment date are the authority; the wait is a function of them.

```{r rebuild}
business_days <- function(from, to, holidays = as.Date(character(0))) {
  vapply(seq_along(from), function(i) {
    if (is.na(from[i]) || is.na(to[i]) || to[i] < from[i]) return(NA_real_)
    if (to[i] == from[i]) return(0)
    d <- seq(from[i] + 1, to[i], by = "day")
    sum(!(format(d, "%u") %in% c("6", "7")) & !(d %in% holidays))
  }, numeric(1))
}

calls$call_date <- as.Date(c("2019-07-01", "2019-07-01", "2019-07-02",
                             "2019-07-02", "2019-07-03",
                             "2019-07-08", "2019-07-08", "2019-07-09", "2019-07-09"))
calls$wait_derived <- business_days(calls$call_date, calls$appointment_date,
                                    holidays = as.Date("2019-07-04"))
calls[, c("clinic", "call_date", "appointment_date", "wait_derived")]
```

The derived column is missing exactly where it should be, and the guard now
passes:

```{r guard-passes}
mysterycall_guard_contaminated_wait(
  data.frame(appointment_date = calls$appointment_date,
             business_days_until_appointment = calls$wait_derived),
  appointment_col = "appointment_date"
)
```

Keep the original column in the source file, labelled as a known-bad historical
artefact, and let the guard be the thing that prevents its use. A defect that is
documented and blocked is safer than one that has been quietly tidied away.

## Where this fits

- `mysterycall_flag_exclusion_discrepancy()` reports the same class of
  discrepancy as a table, for adjudication rather than refusal.
- `mysterycall_flag_included_na_appointments()` covers the mirror-image problem:
  rows marked as included that have no appointment date.
- The `data-quality` vignette covers upstream validation of phones, names and
  joins, before any outcome variable exists.
