---
title: "Visualizing Provider Geographic Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing Provider Geographic Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The `mysterycall` package provides a family of mapping functions for visualizing
physician geographic data across the United States. These functions cover
interactive Leaflet maps centered on ACOG district boundaries, static ggplot2
maps of Hospital Referral Regions (HRRs), and overlap maps that combine
drive-time isochrones with Census block groups. Use them to communicate where
providers practice, how they are distributed across professional districts, and
which communities fall within a given drive-time catchment area.

## Base map

`map_create_base()` returns a [leaflet](https://rstudio.github.io/leaflet/) map
object pre-configured for US provider mapping. It includes two tile-layer
options — **CartoDB Voyager** (a clean, labeled street map) and **Stadia Toner
Lite** (a minimal black-and-white style) — selectable via a layer-control
widget in the top-right corner. A scale bar is added at the bottom-left. An
optional `title` argument accepts an HTML string that is placed in a control
panel at the top-left of the map.

```{r base-map, eval=FALSE}
library(mysterycall)
library(leaflet)

# Default view: continental United States, zoom 4
map_create_base()

# With a custom HTML title
map_create_base(title = "<strong>Ob-Gyn Provider Map</strong>")

# Zoom in on a specific region
map_create_base(
  title = "Mountain West",
  lat   = 39.5,
  lng   = -111.0,
  zoom  = 6
)
```

The returned leaflet object can be piped into any additional `leaflet::add*()`
calls to layer physician points, isochrone polygons, or district boundaries on
top of the base tiles.

## ACOG district boundaries

The American College of Obstetricians and Gynecologists (ACOG) divides the
United States into eleven geographic districts that govern regional governance
and continuing-education activities. `map_create_acog_districts_sf()` joins
the packaged `acog_districts` lookup table to Natural Earth state geometries and
returns a dissolved `sf` polygon object — one row per district — that is ready
for mapping or spatial joins.

```{r acog-districts, eval=FALSE}
library(mysterycall)

# Build the sf object (requires rnaturalearth)
districts_sf <- map_create_acog_districts_sf()

# Inspect the result
class(districts_sf)
#> [1] "sf"         "tbl_df"     "tbl"        "data.frame"

names(districts_sf)
#> [1] "ACOG_District"      "Subregion"          "States"
#> [4] "State_Abbreviations" "geometry"

print(districts_sf[, c("ACOG_District", "States")], n = 5)
```

The bundled `acog_districts` data frame contains the underlying state-to-district
mapping and can be inspected without building the sf object:

```{r acog-data, eval=TRUE}
library(mysterycall)

data(acog_districts)
head(acog_districts)
```

The table below summarizes how many states each district covers:

```{r acog-summary-table, eval=TRUE}
data(acog_districts)
dist_summary <- aggregate(State ~ ACOG_District, data = acog_districts, FUN = length)
names(dist_summary)[2] <- "n_states"
dist_summary <- dist_summary[order(dist_summary$ACOG_District), ]
knitr::kable(
  dist_summary,
  col.names = c("ACOG District", "States"),
  caption   = "Number of US states (and territories) in each ACOG district.",
  row.names = FALSE
)
```

```{r acog-district-bar, eval=TRUE, fig.width=6, fig.height=3.8, fig.cap="State count per ACOG district. Districts vary from 2 (District XI — Puerto Rico / US territories) to 8 states."}
data(acog_districts)
dist_counts <- aggregate(State ~ ACOG_District, data = acog_districts, FUN = length)
names(dist_counts)[2] <- "n_states"
dist_counts$ACOG_District <- factor(
  dist_counts$ACOG_District,
  levels = dist_counts$ACOG_District[order(dist_counts$n_states)]
)

ggplot2::ggplot(dist_counts,
    ggplot2::aes(x = n_states, y = ACOG_District, fill = n_states)) +
  ggplot2::geom_col(width = 0.7) +
  ggplot2::geom_text(ggplot2::aes(label = n_states), hjust = -0.2, size = 3.2) +
  ggplot2::scale_fill_gradient(low = "#c6dbef", high = "#2171b5", guide = "none") +
  ggplot2::scale_x_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
  ggplot2::labs(x = "Number of states", y = NULL) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(panel.grid.major.y = ggplot2::element_blank())
```

## Physician dot map

`map_create_physician_dot()` builds an interactive Leaflet dot map from a data
frame that contains `lat`, `long`, `name`, and `ACOG_District` columns. It
layers ACOG district boundaries in red, colors each physician dot according to
their district using the viridis `"magma"` palette (configurable via
`color_palette`), adds a small random jitter to overlapping points, and saves
the result as a self-contained HTML file plus a PNG screenshot to `output_dir`.

```{r physician-dot, eval=FALSE}
library(mysterycall)

# Synthetic provider data — replace with your own data frame
physician_data <- data.frame(
  long          = c(-95.363, -97.743, -98.494, -96.900, -95.370,
                    -87.623, -122.431, -73.935, -84.388, -104.990),
  lat           = c( 29.763,  30.267,  29.424,  32.779,  29.752,
                     41.883,  37.774,  40.730,  33.749,  39.739),
  name          = paste("Physician", 1:10),
  ACOG_District = c(
    "District VII", "District VII", "District VII", "District VII", "District VII",
    "District IV",  "District IX",  "District II",  "District IV",  "District VIII"
  )
)

# Create the dot map and save HTML + PNG to a temporary directory
map_create_physician_dot(
  physician_data = physician_data,
  jitter_range   = 0.05,
  color_palette  = "magma",
  popup_var      = "name",
  output_dir     = tempdir()
)
```

The function invisibly returns the leaflet map object so you can inspect or
extend it. Requires the `leaflet`, `viridis`, `htmlwidgets`, and `webshot`
packages; install PhantomJS with `webshot::install_phantomjs()` before the
first use.

## Hospital Referral Regions

Hospital Referral Regions (HRRs) are the 306 geographic units defined by the
Dartmouth Atlas of Health Care to represent regional health care markets for
tertiary services. `hrr()` downloads the official shapefile (~8 MB) from the
Dartmouth Atlas on first use, caches it in the user's R cache directory, and
returns an `sf` object in CRS 4326. By default Hawaii and Alaska are excluded;
set `remove_HI_AK = FALSE` to retain all regions.

```{r hrr-sf, eval=FALSE}
library(mysterycall)
library(ggplot2)

hrr_sf <- hrr(remove_HI_AK = TRUE)

# Quick look at the object
glimpse(hrr_sf)
```

The returned sf object can be passed directly to `ggplot2::geom_sf()` for
static publication-ready maps:

```{r hrr-map, eval=FALSE}
ggplot(hrr_sf) +
  geom_sf(fill = "#f0f0f0", color = "#999999", linewidth = 0.15) +
  theme_void() +
  labs(title = "Hospital Referral Regions — Continental United States")
```

`hrr_generate_maps()` extends this by overlaying a physician sf object onto a
hexagonal grid, producing a count-per-cell choropleth with Alaska, Hawaii, and
Puerto Rico insets. The result is saved as both TIFF and PNG at journal-ready
resolution (600 dpi by default).

```{r hrr-hex, eval=FALSE}
# physician_sf must be an sf object with point geometries
hrr_generate_maps(
  physician_sf = physician_sf,
  trait_map    = "obgyn",
  honey_map    = "all",
  output_dir   = tempdir(),
  dpi          = 300,
  width        = 7,
  height       = 5
)
```

## Block group overlap maps

After running `create_isochrones_for_dataframe()` and
`calculate_intersection_overlap_and_save()`, you will have two sf objects: one
with Census block group polygons (containing an `overlap` column — the
proportion of each block group within the isochrone) and one with drive-time
isochrone polygons (containing a `drive_time` column). Pass both to
`map_create_block_group_overlap()` to produce an interactive Leaflet map that
shades block groups by their overlap proportion and draws each isochrone
(30 / 60 / 120 / 180 minutes) in a distinct color. The map is exported as an
HTML file and a PNG screenshot to `output_dir`.

```{r block-group-overlap, eval=FALSE}
library(mysterycall)

# bg_data and isochrones_data are sf objects produced by the isochrone workflow
map_create_block_group_overlap(
  bg_data        = bg_data,
  isochrones_data = isochrones_data,
  output_dir     = tempdir()
)
```

For details on generating `bg_data` and `isochrones_data`, see the
`vignette("create_isochrones", package = "mysterycall")` vignette.
