---
title: "A matched-pair mystery-caller analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A matched-pair mystery-caller analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4, fig.align = "center",
  warning = FALSE, message = FALSE
)
```

This vignette walks a complete matched-pair mystery-caller analysis end to end,
using the tools added for auditing healthcare access. The **matched-pair
design** — the same practice called under two insurance scenarios — is the
paradigm the package is built around: pairing removes each practice's baseline
generosity, so a difference between scenarios is not confounded by *which*
practices happened to be called.

```{r}
library(mysterycall)
```

## 1. A simulated study

We simulate 140 practices, each called twice (Commercial and Medicaid), with a
per-practice random effect so the two calls to one practice are correlated. The
data are in **long** form: one row per practice × scenario.

```{r simulate}
set.seed(2026)
np <- 140L

subspecialty <- sample(c("General", "Pediatric", "Rhinology", "Laryngology"),
                       np, TRUE, prob = c(.45, .25, .18, .12))
region <- sample(c("Northeast", "South", "Midwest", "West"), np, TRUE)
caller <- sample(paste0("RA", 1:5), np, TRUE)
u      <- rnorm(np, 0, 0.7)                              # practice generosity
cbsa   <- ifelse(runif(np) < 0.72,
                 sprintf("C%05d", sample(10000:40000, np, TRUE)), NA_character_)
cfips  <- sprintf("%05d", sample(1001:56045, np, TRUE))

d <- data.frame(
  practice = rep(seq_len(np), times = 2L),
  scenario = rep(c("Commercial", "Medicaid"), each = np),
  stringsAsFactors = FALSE
)
d$subspecialty <- subspecialty[d$practice]
d$region       <- region[d$practice]
d$caller       <- caller[d$practice]
d$cbsa_code    <- cbsa[d$practice]
d$county_fips  <- cfips[d$practice]

medicaid <- d$scenario == "Medicaid"
sub_off  <- c(General = 0, Pediatric = -0.3, Rhinology = 0.2, Laryngology = 0.1)
eta      <- 1.1 - 1.0 * medicaid + u[d$practice] + sub_off[d$subspecialty]

d$office_answered     <- runif(nrow(d)) < 0.94
d$appointment_offered <- d$office_answered & (runif(nrow(d)) < plogis(eta))
d$taking_new_patients <- ifelse(d$appointment_offered | runif(nrow(d)) < 0.6,
                                "Yes", "No")

sub_wait <- c(General = 0, Pediatric = 0.15, Rhinology = 0.1, Laryngology = 0.2)
mu_wait  <- exp(2.4 + 0.45 * medicaid + sub_wait[d$subspecialty])
d$wait_days_business <- ifelse(
  d$appointment_offered, rnbinom(nrow(d), mu = mu_wait, size = 2), NA_real_)

d$appointment_outcome <- ifelse(
  d$appointment_offered,
  sample(c("With sampled physician", "With different physician"),
         nrow(d), TRUE, c(.8, .2)),
  "No appointment offered")

# a `complete` flag: a handful of calls never reached a determination
d$complete <- d$office_answered | runif(nrow(d)) < 0.5

head(d)
```

To make the data-integrity tools earn their keep, we deliberately introduce two
kinds of contradiction that creep into real call logs:

```{r seed-errors}
# (a) a wait time recorded even though no appointment was offered
d$appointment_offered[5]  <- FALSE
d$appointment_outcome[5]  <- "No appointment offered"
d$wait_days_business[5]   <- 14

# (b) the offered flag disagrees with the authoritative outcome
d$appointment_offered[8]  <- FALSE                       # says "not offered" ...
d$appointment_outcome[8]  <- "With sampled physician"    # ... but an appt was made
```

## 2. A clustering key for mixed models

The mixed-model fitters need a `random_intercept` column, but our geography is
split across a metro code that only some practices have and a county code for
the rest. `mysterycall_cluster_id()` coalesces them, giving any still-missing row
its own singleton so it never silently collapses into one giant cluster.

```{r cluster}
d$cluster <- mysterycall_cluster_id(d, c("cbsa_code", "county_fips"))
head(d$cluster)
```

## 3. Data integrity

`mysterycall_check_consistency()` runs a battery of cross-field rules and returns
a single priority-sorted worklist. The defaults include `WAIT_NO_OFFER`, which
catches contradiction (a) above.

```{r consistency}
issues <- mysterycall_check_consistency(d, id_col = "practice")
issues
```

`mysterycall_reconcile_offer_outcome()` finds — and here fixes — rows where the
binary offer flag contradicts the granular outcome (contradiction (b)), treating
the detailed outcome as authoritative and clearing any stray wait/date when a row
flips to "not offered".

```{r reconcile}
rec <- mysterycall_reconcile_offer_outcome(
  d,
  flag_col           = "appointment_offered",
  outcome_col        = "appointment_outcome",
  offered_values     = c("With sampled physician", "With different physician"),
  not_offered_values = "No appointment offered",
  dependent_cols     = "wait_days_business",
  action             = "fix",
  id_col             = "practice"
)
rec
d <- rec$data          # continue with the reconciled data
```

## 4. The access cascade

Simulated-patient studies report a *cascade* of intermediate access measures.
`mysterycall_access_cascade()` turns an ordered set of stage definitions into a
count / denominator / percent table with Wilson intervals, and an optional funnel
figure.

```{r cascade}
cascade <- mysterycall_access_cascade(d, list(
  mysterycall_cascade_stage("Reached a live office", "office_answered", TRUE),
  mysterycall_cascade_stage("Accepting new patients", "taking_new_patients", "Yes"),
  mysterycall_cascade_stage("Appointment offered", "appointment_offered", TRUE),
  mysterycall_cascade_stage("With the sampled physician", "appointment_outcome",
                            "With sampled physician")
), plot = TRUE)
cascade
```

```{r cascade-plot, fig.height = 3.4}
cascade$plot
```

## 5. Bounds under non-response

Not every call completes. Rather than quietly condition the offer rate on the
completed calls, `mysterycall_outcome_bounds()` reports the assumption-free
Manski interval: assign every incomplete call first to failure, then to success.

```{r bounds}
mysterycall_outcome_bounds(d, "appointment_offered", observed = "complete")
```

## 6. Time to an appointment

For a single-contact design the correct time-to-event primitive is the empirical
cumulative proportion offered by business-day *t* — **not** a Kaplan–Meier curve,
which would censor the never-offered calls and imply follow-up the design does
not have. Each curve plateaus at that scenario's offer rate.

```{r curve, fig.height = 3.6}
curve <- mysterycall_cumulative_access_curve(
  d, time_col = "wait_days_business", offered_col = "appointment_offered",
  group_col = "scenario", horizon = 60, plot = TRUE)
curve$plateau
curve$plot
```

## 7. The matched comparison

Now the heart of the design. Only practices called under **both** scenarios
contribute, and for the binary acceptance outcome only the *discordant* practices
(a different answer to the two callers) carry information — so the effective
sample size is the discordant count. `mysterycall_paired_acceptance_mcnemar()`
runs the exact McNemar test.

```{r mcnemar}
d$accepted <- d$appointment_offered
mysterycall_paired_acceptance_mcnemar(
  d, id_col = "practice", scenario_col = "scenario", outcome_col = "accepted")
```

The continuous analogue pairs the within-practice wait differences:

```{r paired-wait}
mysterycall_paired_wait_within_practice(
  d, id_col = "practice", scenario_col = "scenario",
  wait_col = "wait_days_business")
```

## 8. Small-sample categorical and rank tests

For the unmatched views, the categorical toolkit auto-selects chi-square or an
exact test by expected cell counts and reports Cramér's V:

```{r categorical}
mysterycall_test_categorical(d, row_var = "scenario", col_var = "appointment_offered")
```

Offer prevalence with Wilson intervals, by scenario:

```{r prevalence}
mysterycall_prevalence_ci(d, var = "appointment_offered", group_var = "scenario")
```

A Cochran–Mantel–Haenszel test holds region constant while comparing scenarios:

```{r cmh}
mysterycall_cmh_test(d, outcome_var = "appointment_offered",
                     group_var = "scenario", strata_var = "region")
```

Wait times are skewed counts, so compare them with ranks — here across the four
subspecialties, among offered calls:

```{r ranks}
offered <- d[d$appointment_offered & !is.na(d$wait_days_business), ]
mysterycall_compare_ranks(offered, outcome_var = "wait_days_business",
                          group_var = "subspecialty")
```

## 9. Call outcomes and guideline concordance

Raw dispositions map to a standard taxonomy via an ordered keyword list
(first match wins):

```{r classify}
raw <- ifelse(d$appointment_offered, "appointment scheduled",
       sample(c("not accepting new medicaid patients", "no openings for months",
                "does not treat this complaint", "left a voicemail"),
              nrow(d), TRUE))
table(mysterycall_classify_call_outcome(raw))
```

For a guideline-concordance study, build a rubric and score each call transcript
against it. Here three binary items on counseling quality:

```{r concordance}
rubric <- mysterycall_concordance_rubric(
  item  = c("explained_wait", "offered_alternative", "gave_callback"),
  label = c("Explained the expected wait", "Offered an alternative site",
            "Gave a call-back number"),
  type  = "binary", correct = TRUE)

set.seed(11)
transcripts <- data.frame(
  explained_wait      = rbinom(80, 1, 0.7) == 1,
  offered_alternative = rbinom(80, 1, 0.4) == 1,
  gave_callback       = rbinom(80, 1, 0.6) == 1
)
mysterycall_score_concordance(transcripts, rubric, output_dir = NA)
```

## 10. Adjusted model and robustness

Fit an adjusted model for the offer, then interrogate it. (We use a base `glm`
here for speed; `mysterycall_logistic_model()` gives the mixed-model version and
the same robustness helpers accept its objects.)

```{r model}
fit <- glm(appointment_offered ~ scenario + subspecialty + region,
           family = binomial, data = d)
round(exp(cbind(OR = coef(fit), confint.default(fit)))["scenarioMedicaid", ], 2)
```

Is subspecialty jointly informative? `mysterycall_joint_test()` runs the joint
likelihood-ratio test, computing the degrees of freedom from the log-likelihoods
(which sidesteps a well-known `glmer`-vs-`glmmTMB` column-naming footgun):

```{r joint}
mysterycall_joint_test(fit, "subspecialty")
```

Does the Medicaid effect hinge on any single caller?
`mysterycall_leave_one_out()` refits with each caller dropped in turn:

```{r loo}
mysterycall_leave_one_out(fit, d, group = "caller", term = "scenarioMedicaid",
                          joint_predictor = "subspecialty")
```

## 11. Power, for the next study

Access audits have a two-part outcome — a binary offer on the full sample and a
wait time only among offered calls — so both parts must be powered.
`mysterycall_twopart_power()` simulates both.

```{r twopart, eval = requireNamespace("MASS", quietly = TRUE)}
mysterycall_twopart_power(
  n_total = c(200, 400), offer_ref = 0.75, offer_trt = 0.55,
  wait_ref = 14, wait_trt = 21, phi = 1.7, n_sim = 60, seed = 1)
```

Any simulation-based power should be sanity-checked for calibration — the
rejection rate under the null must sit near α. `mysterycall_type_i_check()` turns
an observed null-rejection count into a verdict:

```{r type-i}
mysterycall_type_i_check(rejections = 4, n_sim = 100, alpha = 0.05)
```

And `mysterycall_find_mde()` bisects any power function to the smallest
detectable effect at a target power — here the smallest Medicaid wait (in days)
detectable at 80% power with 400 calls:

```{r mde, eval = requireNamespace("MASS", quietly = TRUE)}
pf <- function(wait_trt) {
  mysterycall_twopart_power(
    n_total = 400, offer_ref = 0.7, offer_trt = 0.7,
    wait_ref = 14, wait_trt = wait_trt, phi = 1.7, n_sim = 30, seed = 1
  )$table$pow_wait
}
mysterycall_find_mde(pf, lower = 14, upper = 26, target_power = 0.80,
                     tol = 0.5, max_iter = 8)$mde
```

When a study oversamples a rare stratum but wants a population-generalizable
effect, `mysterycall_marginal_power()` powers the post-stratification-weighted
marginal estimand (it needs `glmmTMB` and `marginaleffects`):

```{r marginal, eval = FALSE}
mysterycall_marginal_power(
  n_subject = c(150, 300),
  cell_means = c(14, 18, 17, 24),      # Commercial/Medicaid x urban/rural waits
  sigma_subject = 0.4, phi = 1.7,
  stratum_sampling = 0.5, pop_stratum = 0.15,
  n_sim = 200, seed = 1)
```

## Recap

One long call log, start to finish: a clustering key, two integrity passes, the
access cascade, non-response bounds, the cumulative-access curve, the matched
McNemar and paired-wait comparisons, small-sample categorical and rank tests,
outcome classification and guideline concordance, an adjusted model with a joint
test and leave-one-out sensitivity, and simulation-based power for the follow-up
study — each a single function call.
