---
title: "Designing a matched-control mystery-caller audit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing a matched-control mystery-caller audit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 3.6, fig.align = "center"
)
```

Some mystery-caller studies are **case–control** by design: a treated group of
practices — private-equity-owned, hospital-acquired, or a specific chain — is
compared against otherwise-similar practices. Two problems have to be solved
*before* any calls are placed:

1. **Fairness.** Treated and control practices differ on things other than the
   exposure (physician age, degree, gender, local market). If controls are drawn
   at random, an apparent access difference may just reflect those imbalances.
2. **Feasibility.** Callers can only place so many calls, and a control is only
   useful if it is genuinely *callable* — ideally in the same state and a short
   drive from its treated counterpart, so the two share a labor market and payer
   mix.

Matching each treated practice to a nearby, comparable control addresses both.
This vignette walks that design end to end — geocoding a roster, building a
propensity-score-matched control cohort inside a geographic caliper, assembling
the paired calling list, and analysing the completed audit — motivated by a
study of private-equity ownership of OB/GYN practices. (Deriving the treated
cohort itself — tying practice names back to an ownership reference — is an
upstream step outside the package.)

```{r}
library(mysterycall)
```

## 1. The treated cohort and a candidate pool

We simulate 60 private-equity-owned ("treated") practices concentrated in three
metros, plus a large pool of independent candidate controls in the same metros.
Each practice carries the covariates we want to balance — years in practice,
physician gender, and degree (MD vs DO) — and a city/state. The treated
practices are drawn slightly younger and more often DO-led, so the arms are *not*
balanced to begin with.

```{r simulate}
set.seed(2026)
metros <- list(
  CO = c("Denver", "Aurora", "Lakewood"),
  TX = c("Dallas", "Irving", "Plano"),
  FL = c("Miami", "Hialeah", "Miami Beach")
)
draw <- function(n, treated) {
  st <- sample(names(metros), n, TRUE)
  data.frame(
    npi   = NA_integer_,
    state = st,
    city  = vapply(st, function(s) sample(metros[[s]], 1), character(1)),
    years = round(rpois(n, if (treated) 11 else 12)),
    female = rbinom(n, 1, if (treated) 0.45 else 0.55),
    md = rbinom(n, 1, if (treated) 0.72 else 0.82),
    stringsAsFactors = FALSE
  )
}
treated    <- draw(60, TRUE);   treated$npi    <- 1:60
candidates <- draw(600, FALSE); candidates$npi <- 1001:1600
head(treated)
```

## 2. Aligning rosters by name

Before matching, you often have to reconcile two rosters that spell the same
practice differently — the ownership list that defines "treated", and the NPPES
or CMS roster the candidate pool comes from. Exact joins fail on `LLC` vs `PC`,
`&` vs `and`, and stray punctuation. `mysterycall_normalize_org_name()` collapses
those variants to a single canonical key:

```{r normalize}
messy <- c("Women's Health Specialists, LLC", "Womens Health Specialists PC",
           "Rocky Mountain OB & GYN, P.A.")
mysterycall_normalize_org_name(messy)
```

In practice you would add the key to both frames and merge on it:

```{r normalize-join}
cohort <- data.frame(name = messy)
roster <- data.frame(name = c("WOMENS HEALTH SPECIALISTS", "ROCKY MOUNTAIN OB AND GYN"),
                     owner = c("PE-backed", "Independent"))
cohort$key <- mysterycall_normalize_org_name(cohort$name)
merge(cohort, roster, by.x = "key", by.y = "name")
```

## 3. Geocode from city + state

A geographic caliper needs coordinates, but the roster only has a city and a
state. `mysterycall_geocode_city_state()` looks them up in the bundled place
table (~32,000 US places) — no network, no API key — enough to seed a caliper or
QC how far apart matched practices sit. For rooftop-accurate coordinates from
full street addresses, use a dedicated geocoding service instead.

```{r geocode}
tg <- mysterycall_geocode_city_state(treated$city, treated$state)
treated$lat <- tg$lat; treated$lon <- tg$lon
cg <- mysterycall_geocode_city_state(candidates$city, candidates$state)
candidates$lat <- cg$lat; candidates$lon <- cg$lon
c(treated_resolved = mean(!is.na(treated$lat)),
  candidate_resolved = mean(!is.na(candidates$lat)))
```

Places the bundled table misses (townships, renamed municipalities) can be
supplied through the `overrides` argument rather than dropped.

## 4. The propensity model and common support

Matching rests on the **propensity score** — the modeled probability of being
treated given the covariates. `mysterycall_build_matched_controls()` fits it
internally, but it is worth inspecting. Here is the same model, fit directly:

```{r ps-model}
ps_df <- rbind(cbind(treatment = 1, treated[, c("years", "female", "md")]),
               cbind(treatment = 0, candidates[, c("years", "female", "md")]))
ps_fit <- glm(treatment ~ years + female + md, binomial, ps_df)
round(coef(summary(ps_fit)), 3)
```

Matching only works where treated and control scores **overlap** (the
common-support assumption). A treated practice with a score no control comes
close to has no honest match. Compare the two score distributions:

```{r overlap}
treated$ps    <- predict(ps_fit, treated, type = "response")
candidates$ps <- predict(ps_fit, candidates, type = "response")
library(ggplot2)
ggplot(mapping = aes(x = ps)) +
  geom_density(data = candidates, aes(fill = "Candidate pool"), alpha = 0.5) +
  geom_density(data = treated, aes(fill = "Treated (PE)"), alpha = 0.5) +
  scale_fill_manual(NULL, values = c("Candidate pool" = "grey70",
                                     "Treated (PE)" = "#08519c")) +
  labs(x = "Propensity score", y = "Density",
       title = "Propensity-score overlap between arms") +
  theme_minimal(base_size = 12) + theme(legend.position = "top")
```

Good overlap here: every treated score is well inside the candidate range.

## 5. Build the matched control cohort

Now the match itself. `mysterycall_build_matched_controls()` fits the propensity
model, then greedily matches each treated practice 1:1 to its nearest control on
the propensity score — within the **same state** and a **25-mile** caliper,
**without replacement** (each control is used at most once). The greedy scheme,
with a hard state + distance constraint, is what makes every control genuinely
callable.

```{r match}
m <- mysterycall_build_matched_controls(
  treated, candidates,
  ps_formula    = treatment ~ years + female + md,
  id_col        = "npi",
  exact         = "state",
  coords        = c("lat", "lon"),
  caliper_miles = 25
)
m
```

```{r pairs}
head(m$pairs)
```

## 6. Did matching work? Balance before and after

The point of matching is to erase the covariate imbalance we built in. Compare
the standardized mean difference (SMD) of each covariate **before** matching
(treated vs. the whole candidate pool) and **after** (treated vs. matched
controls). An SMD below about 0.1 is the usual "well-balanced" rule of thumb.

```{r balance}
smd <- function(a, b) {
  (mean(a) - mean(b)) / sqrt((var(a) + var(b)) / 2)
}
before <- vapply(c("years", "female", "md"),
                 function(v) smd(treated[[v]], candidates[[v]]), numeric(1))
after  <- setNames(m$balance$smd, m$balance$covariate)
round(data.frame(before_matching = before, after_matching = after[names(before)]), 3)
```

Every covariate moves toward zero after matching.

## 7. How tight a caliper? A sensitivity sweep

The caliper trades geographic closeness against how many treated practices find
a match. Tighten it and controls are nearer but more treated practices go
unmatched. Sweep a few values:

```{r caliper-sweep}
sweep <- sapply(c(5, 10, 25, 50, 100), function(cm) {
  mm <- mysterycall_build_matched_controls(
    treated, candidates, treatment ~ years + female + md,
    id_col = "npi", exact = "state", coords = c("lat", "lon"),
    caliper_miles = cm)
  c(caliper_miles = cm, matched = mm$n_matched,
    median_dist = round(stats::median(mm$pairs$distance_miles, na.rm = TRUE), 1))
})
t(sweep)
```

Report the caliper you chose and this trade-off in the methods; it is a
reviewer's first question.

## 8. The matched calling list

Stack the two arms into one long calling list, tagged by matched pair — what the
callers work from. (`mysterycall_export_gsheet_caller_list()` writes this out in
a Google-Sheets-ready format.)

```{r callist}
pe   <- m$matched_treated;  pe$arm   <- "PE";      pe$pair_id   <- m$pairs$pair_id
ctl  <- m$matched_controls; ctl$arm  <- "Control"; ctl$pair_id  <- m$pairs$pair_id
cols <- c("npi", "state", "arm", "pair_id")
call_list <- rbind(pe[, cols], ctl[, cols])
table(call_list$arm)
```

## 9. Analyse the paired audit

Suppose the calls are complete. We simulate an offer and a wait time for each
practice, with PE practices a little less likely to accept and a little slower:

```{r outcomes}
is_pe <- call_list$arm == "PE"
call_list$accepted  <- rbinom(nrow(call_list), 1, ifelse(is_pe, 0.62, 0.78))
call_list$wait_days <- ifelse(
  call_list$accepted == 1,
  rnbinom(nrow(call_list), mu = ifelse(is_pe, 24, 18), size = 3), NA)
```

Because the design is matched, the comparison is *within pair*, and only the
**discordant** pairs (one arm accepted, the other did not) carry information
about the exposure. The same within-cluster tools used for insurance-scenario
studies apply, with the matched pair as the unit and the arm as the "scenario".
Acceptance, via the exact McNemar test:

```{r mcnemar}
mysterycall_paired_acceptance_mcnemar(
  call_list, id_col = "pair_id", scenario_col = "arm", outcome_col = "accepted")
```

The `discordant`, `disc_favor_a`/`disc_favor_b`, and `odds_ratio` columns show
how many pairs split each way and the direction of the effect; `mde_or_power` is
the smallest odds ratio this many discordant pairs could detect. Wait time, via
the within-pair paired comparison (mean difference with a paired-t CI, plus the
Wilcoxon signed-rank p-value):

```{r paired-wait}
mysterycall_paired_wait_within_practice(
  call_list, id_col = "pair_id", scenario_col = "arm", wait_col = "wait_days")
```

## 10. Power for the next matched study

Sizing a follow-up uses the two-part power tools, treating the arm as the
exposure — a binary offer on every call plus a wait time only among offers:

```{r power, eval = requireNamespace("MASS", quietly = TRUE)}
mysterycall_twopart_power(
  n_total = c(120, 200), offer_ref = 0.78, offer_trt = 0.62,
  wait_ref = 18, wait_trt = 24, phi = 3, n_sim = 50, seed = 1)
```

## Assumptions and limitations

- **Matching balances only *observed* covariates.** Unmeasured differences
  between PE-owned and independent practices can still confound the comparison;
  matching is not randomization.
- **Greedy, not optimal.** Matches are made one treated unit at a time in a fixed
  order, so the total distance is not globally minimized. For this design — where
  a callable, in-caliper control matters more than the global optimum — greedy is
  the right trade-off; when it is not, an optimal matcher (e.g. the \pkg{MatchIt}
  package) is the alternative.
- **City-level geocoding.** `mysterycall_geocode_city_state()` places every
  practice at its city centroid, so within-city pairs read as zero miles apart.
  That is fine for a coarse caliper; a study reporting exact separations should
  geocode full street addresses.
- **Report the unmatched.** Treated practices with no in-caliper control are
  dropped from the analysis; state how many (`m$n_unmatched`) and whether they
  differ systematically, since that shapes who the study generalizes to.

## Recap

From a treated cohort and a candidate pool to a fair, callable comparison and its
analysis: reconcile rosters by normalized name, geocode, check propensity-score
overlap, build a matched control group inside a geographic caliper, confirm
balance improved, choose the caliper deliberately, assemble the paired calling
list, and analyse the completed audit with the matched McNemar and paired-wait
tests — then power the next one.
