---
title: "Designing a matched-control mystery-caller audit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing a matched-control mystery-caller audit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE
)
```

Some mystery-caller studies are **case–control** by design: a treated group of
practices — private-equity-owned, hospital-acquired, or a specific chain — is
compared against otherwise-similar practices. To keep the comparison fair and
the calling burden bounded, each treated practice is matched to a nearby,
comparable control *before* any calls are placed. This vignette builds that
matched cohort and then analyses the resulting paired audit, motivated by a
study of private-equity ownership of OB/GYN practices.

```{r}
library(mysterycall)
```

## 1. The treated cohort and a candidate pool

We simulate 60 private-equity-owned ("treated") practices concentrated in three
metros, plus a large pool of independent candidate controls in the same metros.
Each practice carries the covariates we want to balance — years in practice,
physician gender, and degree — and a city/state.

```{r simulate}
set.seed(2026)
metros <- list(
  CO = c("Denver", "Aurora", "Lakewood"),
  TX = c("Dallas", "Irving", "Plano"),
  FL = c("Miami", "Hialeah", "Miami Beach")
)
draw <- function(n, treated) {
  st <- sample(names(metros), n, TRUE)
  data.frame(
    npi   = NA_integer_,
    state = st,
    city  = vapply(st, function(s) sample(metros[[s]], 1), character(1)),
    years = round(rpois(n, if (treated) 11 else 12)),
    female = rbinom(n, 1, if (treated) 0.45 else 0.55),
    md = rbinom(n, 1, 0.8),
    stringsAsFactors = FALSE
  )
}
treated    <- draw(60, TRUE);   treated$npi    <- 1:60
candidates <- draw(600, FALSE); candidates$npi <- 1001:1600
head(treated)
```

## 2. Geocode from city + state

A caliper needs coordinates, but the roster only has a city and a state.
`mysterycall_geocode_city_state()` looks them up in the bundled place table — no
network, no API key — enough to seed a distance caliper.

```{r geocode}
tg <- mysterycall_geocode_city_state(treated$city, treated$state)
treated$lat <- tg$lat; treated$lon <- tg$lon
cg <- mysterycall_geocode_city_state(candidates$city, candidates$state)
candidates$lat <- cg$lat; candidates$lon <- cg$lon
head(tg)
```

(If instead you needed to align two rosters that spell practice names
differently, `mysterycall_normalize_org_name()` builds a join key:
`mysterycall_normalize_org_name(c("Women's Health Specialists, LLC", "Womens Health Specialists PC"))`
collapses both to `"WOMENS HEALTH SPECIALISTS"`.)

## 3. Build the matched control cohort

`mysterycall_build_matched_controls()` fits a propensity model (treated vs. the
candidate pool), then greedily matches each treated practice 1:1 to its nearest
control on the propensity score — within the **same state** and a **25-mile**
caliper, without replacement.

```{r match}
m <- mysterycall_build_matched_controls(
  treated, candidates,
  ps_formula    = treatment ~ years + female + md,
  id_col        = "npi",
  exact         = "state",
  coords        = c("lat", "lon"),
  caliper_miles = 25
)
m
```

The `balance` table shows the covariates are well-aligned between arms after
matching (small standardized mean differences), and every control is a distinct
practice within 25 miles of its PE match:

```{r pairs}
head(m$pairs)
```

## 4. The matched calling list

Stack the two arms into one long calling list, tagged by matched pair. This is
what the callers would work from.

```{r callist}
pe   <- m$matched_treated;  pe$arm   <- "PE";      pe$pair_id   <- m$pairs$pair_id
ctl  <- m$matched_controls; ctl$arm  <- "Control"; ctl$pair_id  <- m$pairs$pair_id
cols <- c("npi", "state", "arm", "pair_id")
call_list <- rbind(pe[, cols], ctl[, cols])
table(call_list$arm)
```

## 5. Analyse the paired audit

Now suppose the calls are complete. We simulate an offer and a wait time for each
practice, with PE practices a little less likely to accept and a little slower:

```{r outcomes}
is_pe <- call_list$arm == "PE"
call_list$accepted  <- rbinom(nrow(call_list), 1, ifelse(is_pe, 0.62, 0.78))
call_list$wait_days <- ifelse(
  call_list$accepted == 1,
  rnbinom(nrow(call_list), mu = ifelse(is_pe, 24, 18), size = 3), NA)
```

Because the design is matched, the comparison is *within pair*. The same
within-cluster tools used for insurance-scenario studies apply here, with the
matched pair as the unit and the arm as the "scenario". Acceptance, via the exact
McNemar test:

```{r mcnemar}
mysterycall_paired_acceptance_mcnemar(
  call_list, id_col = "pair_id", scenario_col = "arm", outcome_col = "accepted")
```

Wait time, via the within-pair paired comparison:

```{r paired-wait}
mysterycall_paired_wait_within_practice(
  call_list, id_col = "pair_id", scenario_col = "arm", wait_col = "wait_days")
```

## 6. Power for the next matched study

Sizing a follow-up matched audit uses the two-part power tools, treating the arm
as the exposure:

```{r power, eval = requireNamespace("MASS", quietly = TRUE)}
mysterycall_twopart_power(
  n_total = c(120, 200), offer_ref = 0.78, offer_trt = 0.62,
  wait_ref = 18, wait_trt = 24, phi = 3, n_sim = 50, seed = 1)
```

## Recap

From a treated cohort and a candidate pool to a fair, callable comparison and its
analysis: geocode the roster, build a propensity-score-matched control group
inside a geographic caliper, assemble the paired calling list, and analyse the
completed audit with the matched McNemar and paired-wait tests — then power the
next one. Deriving the treated cohort itself (e.g. tying practice names back to a
private-equity ownership reference) is an upstream step outside the package.
