---
title: "Mystery Caller Workflow: From Raw Data to Manuscript"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mystery Caller Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Overview

Mystery-caller (secret-shopper) studies measure disparities in healthcare access by
having trained callers pose as patients requesting appointments under different insurance
types. Each physician in the sampling frame is called once per insurance type; the
caller records whether an appointment was offered and, if so, how many business days
until the earliest available slot. The `mysterycall` package provides a structured
pipeline from raw call logs to manuscript-ready statistics, covering quality control,
deduplication, descriptive summaries, regression modeling, and sensitivity analyses.

## Simulated Data

We construct a 200-row data frame representing 100 physician offices, each called
twice — once under Medicaid and once under Blue Cross/Blue Shield.

```{r simulate, eval=TRUE}
suppressMessages({
  library(mysterycall)

  set.seed(42)
  n_phy  <- 100L
  ids    <- sprintf("%03d", seq_len(n_phy))
  phones <- paste0("555-", sprintf("%04d", seq_len(n_phy)))

  raw_df <- data.frame(
    id_number             = rep(ids, each = 2L),
    phone                 = rep(phones, each = 2L),
    physician_information = paste0("Dr. Physician ", rep(ids, each = 2L)),
    insurance             = rep(c("Medicaid", "Blue Cross/Blue Shield"), n_phy),
    stringsAsFactors      = FALSE
  )

  raw_df$reason_for_exclusions <- sample(
    c("Able to contact", "Physician not available", "Number disconnected"),
    200L, replace = TRUE, prob = c(0.70, 0.20, 0.10)
  )

  # Appointment wait times only for successfully-contacted rows
  raw_df$business_days_until_appointment <- ifelse(
    raw_df$reason_for_exclusions == "Able to contact",
    rpois(200L, lambda = 14L) + 1L,
    NA_integer_
  )

  # Medicaid acceptance column: realistic for Medicaid rows, boilerplate for BCBS
  medicaid_vals <- sample(
    c(
      "Yes they accept Medicaid",
      "No",
      "No answer, unable to determine if they accept Medicaid."
    ),
    200L, replace = TRUE, prob = c(0.55, 0.30, 0.15)
  )
  raw_df$does_the_physician_accept_medicaid <- ifelse(
    raw_df$insurance == "Medicaid",
    medicaid_vals,
    "NA as this was a Blue Cross/Blue Shield call."
  )

  raw_df$scenario <- sample(
    c("HIP scenario", "SHOULDER scenario", "KNEE scenario"),
    200L, replace = TRUE
  )
  raw_df$gender <- sample(c("Male", "Female"), 200L, replace = TRUE)
  raw_df$state  <- sample(
    c(
      "Colorado", "Texas", "California", "New York", "Florida",
      "Illinois", "Pennsylvania", "Ohio", "Georgia", "Michigan"
    ),
    200L, replace = TRUE
  )
  raw_df$notes <- NA_character_
})

dim(raw_df)
head(raw_df[, c("id_number", "phone", "insurance",
                "reason_for_exclusions", "business_days_until_appointment")])
```

## Quality Control

Two QC checks are run on the raw data.  `mysterycall_flag_repeat_physicians()`
identifies physician records appearing more than the expected number of times.
`mysterycall_flag_exclusion_discrepancy()` flags rows that are marked as excluded
but nonetheless carry a non-negative appointment wait time — a logical contradiction
indicating a data-entry error.

```{r qc, eval=TRUE}
suppressMessages({
  repeat_flag <- mysterycall_flag_repeat_physicians(
    raw_df,
    output_dir = NA
  )

  disc_flag <- mysterycall_flag_exclusion_discrepancy(
    raw_df,
    output_dir = NA
  )
})

repeat_flag
disc_flag
```

Both checks pass: no physician exceeds the call-count threshold and no excluded row
carries a valid wait time.

## Deduplication

`mysterycall_dedup_by_insurance()` removes duplicate rows so each
phone x insurance x physician combination appears exactly once.

```{r dedup, eval=TRUE}
df <- suppressMessages(
  mysterycall_dedup_by_insurance(raw_df, output_dir = NA)
)
nrow(df)
```

## Clean Medicaid Column

`mysterycall_clean_medicaid_col()` recodes ambiguous or non-informative responses
to `NA` and appends a binary 0/1 numeric indicator.

```{r clean-medicaid, eval=TRUE}
df <- suppressMessages(
  mysterycall_clean_medicaid_col(df)
)
table(df$cleaned_does_the_physician_accept_medicaid, useNA = "ifany")
```

## Sample Demographics

`mysterycall_sample_demographics()` computes total call counts, insurance-group
physician counts, and geographic coverage, then returns a manuscript-ready sentence.

```{r demographics, eval=TRUE}
demo <- suppressMessages(
  mysterycall_sample_demographics(df, output_dir = NA)
)
cat(demo$summary_sentence, "\n")
```

## Scenario Counts

`mysterycall_scenario_summary()` tabulates calls by scenario and produces a
descriptive sentence ready for insertion into the Participants section.

```{r scenario, eval=TRUE}
scen <- suppressMessages(
  mysterycall_scenario_summary(
    df,
    scenario_levels = c(
      hip      = "HIP scenario",
      shoulder = "SHOULDER scenario",
      knee     = "KNEE scenario"
    ),
    contact_col = "reason_for_exclusions",
    output_dir  = NA
  )
)
print(scen$counts)
cat(scen$sentence, "\n")
```

## Acceptance Rates

`mysterycall_insurance_acceptance_rates()` computes Medicaid and Blue Cross/Blue Shield
acceptance rates using distinct-physician numerators and denominators that exclude
unreachable physician offices.

```{r acceptance, eval=TRUE}
rates <- suppressMessages(
  mysterycall_insurance_acceptance_rates(df, output_dir = NA)
)
cat(rates$paragraph, "\n")
```

## Wait Times

`mysterycall_wait_time_by_group()` provides median (IQR) wait times per group.
`mysterycall_wait_time_sentence()` additionally fits a Poisson GLM to obtain
p-values for each group versus the reference and assembles a full paragraph.

```{r wait-times, eval=TRUE}
wt_grp <- suppressMessages(
  mysterycall_wait_time_by_group(
    df,
    group_col  = "insurance",
    output_dir = NA
  )
)
print(wt_grp)

wt_sent <- suppressMessages(
  mysterycall_wait_time_sentence(df, group_col = "insurance")
)
cat(wt_sent$sentence, "\n")
```

## Poisson Model

`mysterycall_simple_poisson()` fits a Poisson GLM and returns an incidence rate ratio
(IRR) table with confidence intervals and a manuscript-ready summary statement.
Poisson regression is preferred over rank tests because the outcome (days until
appointment) is a count variable.

```{r poisson, eval=TRUE}
poisson_res <- suppressMessages(
  mysterycall_simple_poisson(
    df,
    outcome        = "business_days_until_appointment",
    group          = "insurance",
    reference      = "Blue Cross/Blue Shield",
    outcome_label  = "business days until appointment",
    use_profile_ci = FALSE
  )
)

knitr::kable(
  poisson_res$irr_table[, c("level", "irr", "ci_lower", "ci_upper", "p_value_fmt")],
  digits    = 3,
  caption   = "Poisson regression incidence rate ratios (reference: Blue Cross/Blue Shield)",
  col.names = c("Group", "IRR", "95% CI Lower", "95% CI Upper", "P-value")
)
```

## Sensitivity Analysis

`mysterycall_sensitivity_both_insurance()` identifies physicians called under both
Medicaid and Blue Cross/Blue Shield and compares their wait times, providing a
paired check of whether the observed acceptance-rate difference holds within physicians.

```{r sensitivity, eval=TRUE}
sens <- suppressMessages(
  mysterycall_sensitivity_both_insurance(df, output_dir = NA)
)
cat(sens$sentence, "\n")
```

## Mapping to Manuscript Sections

The functions used in this vignette map directly to standard manuscript sections.
`mysterycall_sample_demographics()` and `mysterycall_scenario_summary()` supply the
**Participants** paragraph of the Methods section. `mysterycall_insurance_acceptance_rates()`
drives the primary **Results** sentence reporting acceptance rates by insurance type.
`mysterycall_wait_time_sentence()` and `mysterycall_simple_poisson()` together produce
the **Statistical Analysis** paragraph and the IRR table reported in the Results.
Finally, `mysterycall_sensitivity_both_insurance()` supplies the **Sensitivity Analysis**
subsection, which compares paired physician records across insurance types to assess
whether the acceptance-rate gap reflects within-physician differences in scheduling
behavior.
