---
title: "Appendix: A Participant Flow That Closes"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    fig_width: 9
    fig_height: 7
description: >
  Validating a STROBE/CONSORT participant flow before drawing it. Why a
  renderer cannot catch a reporting error, what mysterycall_flow_spec()
  checks, and how to put that check in a test suite.
vignette: >
  %\VignetteIndexEntry{Appendix: A Participant Flow That Closes}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(mysterycall)
```

## The problem a renderer cannot see

Every R package for participant-flow diagrams generates the picture from data:
`flowchart`, `consort`, `ggconsort`, `dtrackr`, `vtree`. That is the right
instinct, because it keeps hand-typed numbers out of the figure. It also stops
there. None of them checks that the boxes close.

A flow whose spine reads 1,154 to 1,106 with a 48 exclusion is right. One that
reads 1,154 to 1,100 with a 48 exclusion is a reporting error, and no renderer
will notice, because a renderer draws what it is handed. The figure will look
professional and be wrong, and the error will be discovered by a reviewer or
not at all.

`mysterycall_strobe_flow()` has exactly this gap: it takes `n_total`,
`n_included` and `n_logistic` from the caller and draws them. The two functions
in this appendix are the layer underneath it. Validate first, render second,
and make the validation reusable.

## Validation is a separate step on purpose

```{r}
spec <- mysterycall_flow_spec(
  spine = c(
    "Abstracts parsed from congress supplements" = 1154,
    "Oral presentation cohort"                   = 1106,
    "Evaluated for publication status"           = 1051
  ),
  exclusions = list(
    "Abstracts parsed from congress supplements" = c("Video presentations" = 48),
    "Oral presentation cohort"                   = c("Adjudication unresolved" = 55)
  ),
  splits = list(
    "Evaluated for publication status" = c("Published" = 170, "Not published" = 881),
    "Not published" = c("No qualifying publication" = 839,
                        "Publication predates the congress" = 42)
  )
)

spec$closed
```

Three things are checked:

1. **The spine closes.** Each step, minus the exclusions leaving it, equals the
   next step. Here 1154 − 48 = 1106, and 1106 − 55 = 1051.
2. **Every split sums to its parent.** 170 + 881 = 1051, and within that,
   839 + 42 = 881. Splits may nest to any depth; a child can itself be split.
3. **Every count is a count.** No negative, fractional, `NA`, or duplicated
   values.

### The failure message names the step

An error that says a diagram is inconsistent somewhere is nearly useless in a
flow with fifteen boxes. This one points at the arithmetic that failed:

```{r, error=TRUE}
mysterycall_flow_spec(
  spine      = c("Screened" = 100, "Analysed" = 90),
  exclusions = list("Screened" = c("Ineligible" = 5))
)
```

All failures are collected rather than reported one at a time, so a spec with
two problems tells you both:

```{r, error=TRUE}
mysterycall_flow_spec(
  spine  = c("Screened" = 100, "Analysed" = 95),
  splits = list("Analysed" = c("A" = 50, "B" = 40))
)
```

## Put it in the test suite

This is the reason validation is not folded into the drawing function. A check
that only runs at render time runs on the author's laptop, once, and a figure
that goes stale six months later goes stale silently. The same call in
`tests/testthat/` fails the build:

```r
test_that("the published flow still matches the analysis dataset", {
  spec <- mysterycall_flow_spec(
    spine = c(
      "Screened"  = nrow(screened),
      "Eligible"  = nrow(eligible),
      "Analysed"  = nrow(analysed)
    ),
    exclusions = list(
      "Screened" = c("Did not meet criteria" = nrow(screened) - nrow(eligible)),
      "Eligible" = c("Incomplete outcome"    = nrow(eligible) - nrow(analysed))
    )
  )
  expect_true(spec$closed)
})
```

Wire the counts to the data rather than typing them, and the test tells you the
manuscript figure disagrees with the dataset the moment it does.

## Drawing

`mysterycall_strobe_diagram()` accepts a validated spec and nothing else:

```{r, error=TRUE}
mysterycall_strobe_diagram(list(spine = c(a = 2, b = 1)))
```

```{r fig-flow, fig.alt="Participant flow diagram: 1,154 abstracts to 1,106 to 1,051, with two exclusion boxes and a nested published/not-published split"}
# Note the print(). The function returns its ggplot invisibly, because its
# common use is writing a file via output_path, where an auto-printed plot
# would be noise. In an R Markdown chunk that means nothing is drawn unless
# you print it explicitly.
print(mysterycall_strobe_diagram(spec, title = "Participant flow"))
```

It is ggplot2 only. There is no Graphviz or htmlwidget path, so writing a PNG
needs no headless browser and the figure rebuilds in a plain CI container.
Pass `output_path` to write the file directly.

### Background

The figure is saved on an opaque white background by default.
`ggplot2::theme_void()` leaves the panel blank, which writes a fully
transparent PNG: invisible on a white page, and wrong the moment the figure is
dropped into a Word manuscript or onto a coloured slide, where whatever sits
behind it shows through the boxes. Pass `background = NA` if a transparent
figure is what you actually want.

## Relationship to the other flow functions

| Function | Use it when |
|---|---|
| `mysterycall_flow_spec()` | You want the arithmetic checked, in a test or before drawing |
| `mysterycall_strobe_diagram()` | You have a validated spec and want a ggplot |
| `mysterycall_strobe_flow()` | Older, simpler entry point; takes counts and draws them without validating |
| `mysterycall_flow_diagram()` | Fixed-shape CONSORT-style diagram for the standard audit pipeline |

New work should prefer the `flow_spec()` and `strobe_diagram()` pair, because
it is the only one of the four that can tell you the numbers are wrong.

## See also

`vignette("reporting-conventions")` for how the package formats the numbers
that end up in these boxes, and `mysterycall_strobe_checklist()` for the
STROBE item that asks for a participant flow in the first place.
