---
title: "Power Analysis and Sample Size for Mystery Caller Studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Power Analysis and Sample Size for Mystery Caller Studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 6,
  fig.height = 4,
  out.width = "95%"
)
```

## Overview

A mystery caller (audit) study requires careful pre-study planning to ensure the
resulting data can answer the research question with adequate statistical
precision. This vignette covers three interrelated planning decisions:

1. **Sampling frame** — How many providers from the directory must be called
   (Cochran finite-population formula)?
2. **Power for a count outcome** — How many providers per insurance arm are
   needed to detect a clinically meaningful difference in wait days (Poisson
   power analysis)?
3. **Power curves** — How does required sample size change across a range of
   effect sizes?

The functions shown here are in the `power analysis` family:

| Function | Purpose |
|---|---|
| `mysterycall_cochran_n()` | Finite-population sample size for proportion outcomes |
| `mysterycall_poisson_power()` | Sample size for Poisson count outcomes (wait days) |
| `mysterycall_equation_figure()` | Power curve plot across a range of IRR values |

```{r load}
library(mysterycall)
```

---

## 1. Background: Two Types of Outcomes

Mystery caller studies typically involve two kinds of outcomes:

| Outcome | Example | Statistical model | Planning tool |
|---|---|---|---|
| **Binary** | Was the call accepted? (yes/no) | Logistic regression | Cochran formula |
| **Count** | How many days until appointment? | Poisson / negative-binomial GLMM | Poisson power formula |

The sample size differs for each.

---

## 2. Cochran Finite-Population Sample Size

### Mathematical basis

When the sampling frame is a known, finite directory of $N$ providers, the
Cochran (1977) formula gives the minimum sample $n$ so that the estimated
**proportion** (e.g. acceptance rate) is within $e$ percentage points with
95 % confidence:

$$
n_0 = \frac{Z^2_{\alpha/2} \cdot p(1-p)}{e^2}
\qquad\longrightarrow\qquad
n = \frac{n_0}{1 + \frac{n_0 - 1}{N}}
$$

Setting $p = 0.5$ (maximum variance) and $Z_{0.025} = 1.96$ this simplifies to:

$$
n = \frac{N}{1 + N e^2}
$$

The finite-population correction factor $\left(1 + \frac{n_0-1}{N}\right)^{-1}$
makes $n < n_0$ whenever the sample represents a meaningful fraction of the
directory.

### Using `mysterycall_cochran_n()`

```{r cochran-basic, echo=TRUE, eval=TRUE}
# 800 OB-GYNs in target region — need acceptance rate within ±5 percentage points
res_800 <- mysterycall_cochran_n(N = 800, margin_of_error = 0.05)
cat(sprintf(
  "N = %d providers, target ME ±%.0f%%: sample n = %d (achieved ME = ±%.1f%%)\n",
  res_800$N, res_800$margin_of_error * 100,
  res_800$n, res_800$effective_margin * 100
))
```

```{r cochran-grid, echo=TRUE, eval=TRUE}
# Show how n changes across directory sizes and margin-of-error targets
grid <- expand.grid(
  N  = c(100, 200, 400, 800, 1600),
  me = c(0.03, 0.05, 0.10)
)

grid$n <- mapply(
  function(N, me) mysterycall_cochran_n(N, me)$n,
  grid$N, grid$me
)

pivot_tbl <- stats::reshape(
  grid,
  idvar = "N", timevar = "me", direction = "wide"
)
names(pivot_tbl) <- c("Directory size (N)", "n at ±3%", "n at ±5%", "n at ±10%")

knitr::kable(
  pivot_tbl,
  caption = "Required sample sizes across directory sizes and precision targets."
)
```

**Key insight:** For moderate-to-large directories (N ≥ 400), the ±5%
target requires roughly 200 providers regardless of how large the directory
grows — the finite-population correction becomes negligible.

---

## 3. Poisson Power Analysis for Count Outcomes

### Mathematical basis

Wait times (business days until appointment) are count data analyzed with a
Poisson generalized linear model (or mixed model when physicians are called
with both insurance types). The parameter of interest is the **incidence rate
ratio** (IRR): how many times longer is the Medicaid wait compared with the
private-insurance wait?

The score-test sample size for comparing two independent Poisson rates
$\lambda_0$ (reference arm) and $\lambda_1 = \text{IRR} \times \lambda_0$
(comparison arm) is:

$$
n = \frac{(z_{\alpha/2} + z_\beta)^2
          \left(\dfrac{1}{\lambda_0} + \dfrac{1}{\lambda_1}\right)}{(\ln \text{IRR})^2}
$$

This gives **providers per arm**. In a paired design (each provider is called
with both insurance types) only $n$ providers are needed in total.

### Worked example

Suppose prior published work suggests physicians give Medicaid callers a wait
of about 1.40 times the BCBS wait. The BCBS mean is 14 business days.
How many providers are needed for 80 % power?

```{r poisson-basic, echo=TRUE, eval=TRUE}
pw <- mysterycall_poisson_power(
  irr       = 1.40,   # Medicaid wait is 40% longer
  lambda_ref = 14,    # BCBS mean = 14 business days
  alpha      = 0.05,
  power      = 0.80,
  both_arms  = TRUE   # paired design
)

cat(sprintf(
  "IRR = %.2f | lambda_ref = %.0f d | lambda_trt = %.0f d\n",
  pw$irr, pw$lambda_ref, pw$lambda_trt
))
cat(sprintf(
  "n per arm = %d | total providers = %d | total calls = %d\n",
  pw$n_per_arm, pw$n_total, pw$n_total_calls
))
```

### Comparing power and effect sizes

```{r poisson-table, echo=TRUE, eval=TRUE}
scenarios <- list(
  list(irr = 1.20, power = 0.80, label = "Small effect (IRR 1.20), 80% power"),
  list(irr = 1.40, power = 0.80, label = "Moderate effect (IRR 1.40), 80% power"),
  list(irr = 1.40, power = 0.90, label = "Moderate effect (IRR 1.40), 90% power"),
  list(irr = 1.60, power = 0.80, label = "Large effect (IRR 1.60), 80% power"),
  list(irr = 2.00, power = 0.80, label = "Very large effect (IRR 2.00), 80% power")
)

pw_rows <- lapply(scenarios, function(s) {
  res <- mysterycall_poisson_power(
    irr        = s$irr,
    lambda_ref = 14,
    power      = s$power,
    both_arms  = TRUE
  )
  data.frame(
    Scenario       = s$label,
    IRR            = s$irr,
    Power          = sprintf("%.0f%%", s$power * 100),
    `n per arm`    = res$n_per_arm,
    `Total providers` = res$n_total,
    `Total calls`  = res$n_total_calls,
    check.names    = FALSE
  )
})

pw_tbl <- do.call(rbind, pw_rows)

knitr::kable(
  pw_tbl,
  caption = paste(
    "Poisson power analysis: providers per arm for a paired design,",
    "lambda_ref = 14 days, alpha = 0.05."
  )
)
```

**Interpretation:**

- Detecting a small IRR (1.20) requires substantially more providers than
  detecting a large one (1.60+).
- The paired design (each provider called with both insurance types) halves
  the total providers compared with an independent-groups design.
- Increasing power from 80 % to 90 % adds roughly 25–35 % more providers.

---

## 4. Clustering Adjustment (ICC)

In practice, providers at the same practice share scheduling staff, so calls
to the same office may be correlated. The design effect inflates the sample
size by:

$$
\text{DEFF} = 1 + (m - 1) \times \rho
$$

where $m$ is the average calls per cluster (office) and $\rho$ is the
intra-class correlation (ICC).

```{r icc, echo=TRUE, eval=TRUE}
# No clustering vs. ICC = 0.10 (moderate) with 3 calls per office on average
pw_no_icc  <- mysterycall_poisson_power(irr = 1.40, lambda_ref = 14)
pw_icc_010 <- mysterycall_poisson_power(irr = 1.40, lambda_ref = 14,
                                         icc = 0.10, calls_per_cluster = 3)
pw_icc_020 <- mysterycall_poisson_power(irr = 1.40, lambda_ref = 14,
                                         icc = 0.20, calls_per_cluster = 3)

icc_tbl <- data.frame(
  ICC            = c("0 (none)", "0.10 (moderate)", "0.20 (high)"),
  DEFF           = c(pw_no_icc$design_effect,
                     pw_icc_010$design_effect,
                     pw_icc_020$design_effect),
  `n per arm`    = c(pw_no_icc$n_per_arm,
                     pw_icc_010$n_per_arm,
                     pw_icc_020$n_per_arm),
  `Total calls`  = c(pw_no_icc$n_total_calls,
                     pw_icc_010$n_total_calls,
                     pw_icc_020$n_total_calls),
  check.names    = FALSE
)

knitr::kable(
  icc_tbl,
  digits  = 2,
  caption = paste(
    "Effect of within-office correlation (ICC) on required sample size.",
    "IRR = 1.40, lambda_ref = 14 d, 80% power, m = 3 calls/office."
  )
)
```

> **Rule of thumb:** An ICC of 0.10 with 3 calls per office inflates the
> required sample by ~20 %. When the ICC is unknown, a sensitivity analysis
> across ICC values (0, 0.05, 0.10, 0.20) is recommended.

---

## 5. Power Curves

A power curve shows how the required sample size per arm changes as a function
of the minimum detectable effect (IRR). Plotting the curve helps justify the
study's sample size to reviewers and IRBs.

```{r power-curve, echo=TRUE, eval=TRUE, fig.cap="Required providers per arm (paired design, 80% power, lambda_ref = 14 d) as a function of the minimum detectable IRR."}
mysterycall_equation_figure(
  lambda0   = 14,
  irr_seq   = seq(1.10, 2.0, by = 0.05),
  alpha     = 0.05,
  power     = 0.80,
  both_arms = TRUE
)
```

```{r power-curve-90, echo=TRUE, eval=TRUE, fig.cap="Power curves comparing 80% and 90% power targets."}
library(ggplot2)

irr_seq <- seq(1.10, 2.0, by = 0.05)

pw_80 <- vapply(irr_seq, function(irr)
  mysterycall_poisson_power(irr, lambda_ref = 14, power = 0.80, both_arms = TRUE)$n_per_arm,
  numeric(1))

pw_90 <- vapply(irr_seq, function(irr)
  mysterycall_poisson_power(irr, lambda_ref = 14, power = 0.90, both_arms = TRUE)$n_per_arm,
  numeric(1))

curve_df <- rbind(
  data.frame(irr = irr_seq, n = pw_80, power = "80%"),
  data.frame(irr = irr_seq, n = pw_90, power = "90%")
)

ggplot(curve_df, aes(x = irr, y = n, color = power, linetype = power)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("80%" = "#2C3E50", "90%" = "#C0392B")) +
  labs(
    title   = "Poisson power curves by target power level",
    subtitle = "Paired design, lambda_ref = 14 days, alpha = 0.05",
    x       = "Minimum detectable IRR",
    y       = "Required providers per arm",
    color   = "Power",
    linetype = "Power"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

---

## 6. Sample Size Across Reference Wait Times

The expected wait time for the reference arm (`lambda_ref`) also affects
required sample size. Studies in different specialties or regions should adapt
this parameter.

```{r lambda-sensitivity, echo=TRUE, eval=TRUE, fig.cap="Required sample size per arm for IRR = 1.40 across different reference wait-time assumptions."}
lambda_vals <- c(5, 7, 10, 14, 21, 30)

n_by_lambda <- vapply(lambda_vals, function(lam)
  mysterycall_poisson_power(irr = 1.40, lambda_ref = lam, power = 0.80,
                             both_arms = TRUE)$n_per_arm,
  numeric(1))

lam_df <- data.frame(lambda_ref = lambda_vals, n_per_arm = n_by_lambda)

ggplot(lam_df, aes(x = lambda_ref, y = n_per_arm)) +
  geom_col(fill = "#2166AC", alpha = 0.8) +
  geom_text(aes(label = n_per_arm), vjust = -0.4, size = 3.5) +
  labs(
    title   = "Sample size sensitivity to reference wait time",
    subtitle = "IRR = 1.40, 80% power, paired design",
    x       = "Reference-arm mean wait time (days)",
    y       = "Required providers per arm"
  ) +
  theme_minimal(base_size = 12)
```

**Interpretation:** As wait times get shorter (fewer days), the Poisson rate
becomes smaller and more variable relative to its mean, so more providers are
needed to detect the same proportional difference.

---

## 7. Combining Both Calculations

A well-designed mystery caller study may need to satisfy *both* constraints:

1. The Cochran sample size ensures acceptable **precision** for the acceptance
   rate proportion.
2. The Poisson power ensures **power** to detect a meaningful IRR on wait days.

Take the **larger** of the two as the study's minimum sample size:

```{r combined, echo=TRUE, eval=TRUE}
n_cochran <- mysterycall_cochran_n(N = 800, margin_of_error = 0.05)$n
n_poisson <- mysterycall_poisson_power(irr = 1.40, lambda_ref = 14,
                                        power = 0.80, both_arms = TRUE)$n_total

cat(sprintf("Cochran (proportion, ±5%% ME):  n = %d\n", n_cochran))
cat(sprintf("Poisson power (IRR 1.40, 80%%): n = %d\n", n_poisson))
cat(sprintf("Required study size:           n = %d\n", max(n_cochran, n_poisson)))
```

---

## 8. Reporting Template

When writing the Methods section of a manuscript, the relevant numbers can be
generated programmatically:

```{r reporting, echo=TRUE, eval=TRUE}
cochran_res <- mysterycall_cochran_n(N = 800, margin_of_error = 0.05)
poisson_res <- mysterycall_poisson_power(irr = 1.40, lambda_ref = 14,
                                          power = 0.80, both_arms = TRUE)

cat(sprintf(paste(
  "Of the %d obstetrician-gynecologists in the study directory,",
  "a minimum sample of %d providers was required for the acceptance-rate",
  "outcome to achieve a margin of error of ±%.0f%% (Cochran 1977).",
  "For the wait-time count outcome, a paired-design Poisson power analysis",
  "(IRR = %.2f, lambda_ref = %.0f d, alpha = 0.05, power = 80%%) indicated",
  "that %d providers would be sufficient (n_total_calls = %d)."),
  cochran_res$N,
  cochran_res$n,
  cochran_res$margin_of_error * 100,
  poisson_res$irr,
  poisson_res$lambda_ref,
  poisson_res$n_total,
  poisson_res$n_total_calls
))
```

---

## 9. References

- Cochran, W. G. (1977). *Sampling Techniques* (3rd ed.). Wiley.
- Zhu, H. (2019). Poisson regression for count data. *Journal of Statistical
  Software*, 2(3). (Score-test sample size derivation.)
- Donner, A., & Klar, N. (2000). *Design and Analysis of Cluster Randomization
  Trials in Health Research*. Arnold. (Design effect / ICC correction.)

---

## See also

- `vignette("statistical-analysis")` — full GLM / GLMM analysis pipeline
- `vignette("workflow-orchestration")` — end-to-end mystery caller workflow
- [`mysterycall_bootstrap_ci()`] — bootstrap confidence intervals for
  acceptance rates
- [`mysterycall_multiple_comparison_adjust()`] — Holm–Bonferroni correction
  for multiple insurance comparisons
