---
title: "Appendix: Statistical Reporting Conventions (SAMPL)"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    df_print: kable
description: >
  How mysterycall prints numbers, and why. The SAMPL guideline items the
  package can enforce mechanically, the two formatters that own the
  convention, and the three checklists that go in a supplementary file.
vignette: >
  %\VignetteIndexEntry{Appendix: Statistical Reporting Conventions (SAMPL)}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE, purl=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(mysterycall)
```

## Why this appendix exists

STROBE says what to report about an observational study's design. SAMPL
(Statistical Analyses and Methods in the Published Literature, Lang and Altman)
says how to report the numbers themselves: give an effect size with a
confidence interval rather than a bare p-value, name the test and why it fits,
give the denominator behind every percentage, report p-values exactly rather
than as inequalities against alpha.

The practical consequence for a mystery-caller manuscript is that reporting

> OR 0.31, 95% CI 0.05 to 1.9

instead of

> p = .21

is a deliberate choice backed by a published guideline, not a way to dodge a
null result. A reviewer who knows the guideline can see that immediately. This
appendix records which of those choices the package makes for you, so that a
methods section can cite them rather than restate them.

## The interval separator is not cosmetic

SAMPL asks that interval endpoints be separated with "to" rather than a hyphen
or a dash. This reads as a typographic nicety until an endpoint goes negative:

```{r}
lo <- -0.45
hi <- -0.12

# what a hyphen produces
sprintf("%.2f-%.2f", lo, hi)

# what the package produces
mysterycall_format_ci(lo, hi)
```

The first string is genuinely ambiguous, and it was what
`mysterycall_multi_model_table()` printed for every linear mixed model, whose
column is labelled `Beta (95% CI)` and whose estimates are routinely negative.
Ratio measures hid the problem, because an odds ratio or an incidence rate
ratio cannot be negative and so always read correctly.

## The two formatters

Every interval and every p-value in the package goes through one of two
exported functions. They exist so the convention lives in one place rather than
being retyped at each call site.

```{r}
mysterycall_format_ci(1.05, 1.57)
mysterycall_format_ci(c(1.05, NA), c(1.57, 2.00))

mysterycall_format_p(c(0.0431, 0.0004, NA))
mysterycall_format_p(0.0431, name = "p")
```

`format_p()` takes `name` because both shapes are needed: the bare value for a
table cell, the prefixed form for prose.

### Changing the house style

The separator is an argument, not a constant, because journals disagree. This
follows the pattern set by `gtsummary`'s JAMA and Lancet journal themes, which
make the same "to" choice and expose it as a `ci.sep` theme setting.

```{r}
# per call
mysterycall_format_ci(1.05, 1.57, sep = " - ")

# for a whole document
withr::with_options(
  list(mysterycall.ci_sep = " to "),
  mysterycall_format_ci(1.05, 1.57)
)
```

Set `options(mysterycall.ci_sep = )` once in a setup chunk and every table,
plot label, and generated sentence follows.

## What the package decides for you

| SAMPL item | How the package handles it |
|---|---|
| Interval endpoints separated with "to" | `mysterycall_format_ci()`, used by every table and sentence builder |
| p-values reported exactly, never as "NS" | `mysterycall_format_p()`; collapses only below `threshold` |
| No coding against alpha | `mysterycall_irr_table()` has `add_significance_stars = FALSE` by default |
| Effect size with an interval, not a bare p | Model tables emit estimate and interval together |
| Denominator behind each percentage | Acceptance-rate helpers print `n/N` alongside the percentage |

Three places deliberately depart from this and are not bugs:

- `mysterycall_forest_plot()` uses AMA house style for p-values, meaning a
  capital `P`, no leading zero, and a two-tier digit rule. That is a different
  published convention rather than a lapse.
- `mysterycall_multiple_comparison()` and `mysterycall_subspecialist_trend()`
  emit significance stars, because their output is a screening display rather
  than a reported estimate.
- Year ranges and wait-time band labels keep a dash. They are category names,
  not estimates, and cannot take a negative endpoint.

## The three checklists

The package ships three fillable reporting checklists, meant to be bound
together into one supplementary file. They do not overlap.

```{r}
sampl <- mysterycall_sampl_checklist()
nrow(sampl)
table(sampl$section)
```

| Checklist | Covers | Items |
|---|---|---|
| `mysterycall_strobe_checklist()` | Observational study design, auto-checked against a fitted model | 16 |
| `mysterycall_crisp_checklist()` | Simulated-patient methodology: caller training, detection, ethics of deception | 20 |
| `mysterycall_sampl_checklist()` | How the numbers themselves are reported | 27 |

The `reported` column is yours to fill with a page or section reference:

```{r}
head(sampl[, c("section", "item")], 4)
```

### A known limitation

Filling `reported` by hand is the weak point. Nothing ties a checklist item to
the code that satisfies it, so the column records a claim rather than a fact.
rOpenSci's [srr](https://github.com/ropensci-review-tools/srr) solves this
properly: standards are tagged in roxygen at the code location that meets them,
and a pre-submit check fails while any remain unaddressed. That is the model to
follow if these items ever need machine verification.

## References

Lang TA, Altman DG. Basic statistical reporting for articles published in
biomedical journals: the "Statistical Analyses and Methods in the Published
Literature" or the SAMPL Guidelines. *International Journal of Nursing
Studies*. 2015;52(1):5-9. <https://doi.org/10.1016/j.ijnurstu.2014.09.006>

von Elm E, Altman DG, Egger M, Pocock SJ, Gotzsche PC, Vandenbroucke JP.
The Strengthening the Reporting of Observational Studies in Epidemiology
(STROBE) statement. *Lancet*. 2007;370(9596):1453-1457.

The SAMPL guidelines are hosted by the EQUATOR Network at
<https://www.equator-network.org/reporting-guidelines/sampl/>.
