---
title: "Search and Process NPI Numbers"
author:
  - Tyler Muffly, MD
output:
  rmarkdown::html_vignette:
    df_print: kable
description: >
  Search the NPI registry from clinician first and last names using the current
  mysterycall interface.
vignette: >
  %\VignetteIndexEntry{Search and Process NPI Numbers}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
library(magrittr)
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  collapse = TRUE,
  eval = eval_vignettes
)

if (eval_vignettes) {
  library(mysterycall)
}
```

# Overview

`mysterycall_search_and_process_npi()` is the name-based entry point for roster creation.
Use it when you already have clinician first and last names from a patient
directory, credentialing list, or manually curated study roster and need to
recover likely NPI matches.

The current exported interface expects a data frame passed through the `data`
argument. That data frame must contain:

- `first`
- `last`

## Minimal input

```{r}
sample_names <- tibble::tibble(
  first = c("Jane", "Maria", "Alex"),
  last = c("Smith", "Garcia", "Johnson")
)

sample_names
```

## Basic search

The simplest call looks like this:

```{r eval=FALSE}
results <- mysterycall_search_and_process_npi(
  data = sample_names,
  notify = FALSE
)
```

By default the function searches for U.S. individual NPIs and filters to common
physician credentials such as `MD` and `DO`.

## Customizing the search

For most projects you will want to set a few operational parameters explicitly,
especially if the roster is large.

```{r eval=FALSE}
results <- mysterycall_search_and_process_npi(
  data = sample_names,
  enumeration_type = "ind",
  limit = 10,
  country_code = "US",
  filter_credentials = c("MD", "DO"),
  notify = FALSE
)
```

## Tracking long runs

Large name-based searches can take time. The function supports progress logs,
chunked saves, and resumable accumulation files.

```{r eval=FALSE}
results <- mysterycall_search_and_process_npi(
  data = sample_names,
  limit = 10,
  dest_dir = "npi_chunks",
  accumulate_path = "npi_results.csv",
  resume = TRUE,
  progress_log = "npi_progress.csv",
  progress_log_format = "csv",
  heartbeat_seconds = 30,
  notify = FALSE
)
```

This is useful when the roster is large enough that you want durable progress
tracking or expect to resume after an interruption.

## What comes back

The returned object is a data frame of candidate NPI matches. Key columns:

| Column | Description |
|--------|-------------|
| `npi` | 10-digit National Provider Identifier (character) |
| `basic_first_name` | First name from NPPES |
| `basic_last_name` | Last name from NPPES |
| `basic_middle_name` | Middle name from NPPES |
| `basic_gender` | `"M"` or `"F"` as reported to NPPES |
| `basic_credential` | Credential string (e.g. `"MD"`, `"DO"`) |
| `addresses_state` | Two-letter state abbreviation of primary practice |
| `addresses_telephone_number` | Primary phone number (unformatted) |
| `taxonomies_desc` | Self-reported specialty taxonomy description |
| `search_term` | The `"first last"` input pair that produced this match |

```{r eval=FALSE}
dplyr::glimpse(results)

results |>
  dplyr::count(search_term, sort = TRUE)
```

## Cleaning the result set

Name-based matches are often one-to-many. A common first cleanup step is to
deduplicate on `npi` and then inspect specialty or geography fields.

```{r eval=FALSE}
results_unique <- results |>
  dplyr::distinct(npi, .keep_all = TRUE)

results_unique |>
  dplyr::count(taxonomies_desc, sort = TRUE)
```

## Validating returned names and phones

After deduplication, run the data quality helpers on the raw NPPES fields before
any downstream joins:

```{r eval=FALSE}
# Parse names returned from NPPES into structured components
parsed_names <- do.call(rbind, lapply(
  paste(results_unique$basic_first_name, results_unique$basic_last_name),
  mysterycall_parse_physician_name
))

low_conf <- parsed_names[parsed_names$confidence == "low", ]
if (nrow(low_conf)) message(nrow(low_conf), " low-confidence name parses — review manually")

# Validate phone numbers against the provider's reported state
phone_checks <- do.call(rbind, Map(
  mysterycall_validate_phone,
  phone_str      = results_unique$addresses_telephone_number,
  practice_state = results_unique$addresses_state
))

table(phone_checks$validity_flag)
```

## Joining to clinician enrichment safely

Use `mysterycall_safe_left_join()` when attaching CMS Care Compare data to catch
NPI type mismatches and coverage gaps:

```{r eval=FALSE}
clinician_data <- mysterycall_get_clinician_data(results_unique)

roster_enriched <- mysterycall_safe_left_join(
  left         = results_unique,
  right        = clinician_data,
  by           = "npi",
  label_left   = "npi_search_results",
  label_right  = "cms_clinician_data",
  min_coverage = 0.80
)
```

## Practical guidance

- Start with a small roster to confirm that your filters are behaving as
  expected.
- Expect ambiguous matches for common names; always inspect `search_term` counts.
- Save progress when running large overnight jobs.
- Deduplicate and validate phones and names before downstream joins.

# Next step

If your project begins from a specialty rather than a name list, see the
taxonomy-search vignette for `mysterycall_search_taxonomy()`. For phone
validation, name parsing, and safe joins see
`vignette("data-quality", package = "mysterycall")`.
