---
title: "Subspecialist density per 100,000 women"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Subspecialist density per 100,000 women}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE,
  fig.width = 7, fig.height = 4.2, fig.align = "center"
)
library(mysterycall)
```

Workforce-density figures — *subspecialists per 100,000 women* — turn raw
provider counts into an access metric a reader can compare across
subspecialties and over time. This vignette walks the full path: **counts +
denominator → density → figure**, with the provenance that keeps the picture
honest.

Two figures share the same computation:

- `mysterycall_subspecialist_trend()` — a multi-year line trend (one line per
  subspecialty), optionally with confidence bands.
- `mysterycall_subspecialist_infographic()` — a two-point (start → end) badge
  infographic.

The density is always derived from the inputs (`count / population * 100000`),
never typed, so a figure cannot disagree with its own numbers.

## 1. The numerator: subspecialist counts

You supply the counts. A long data frame is the most flexible form — one row
per subspecialty per year:

```{r counts}
counts <- data.frame(
  subspecialty = rep(c("Gynecologic Oncology", "Maternal-Fetal Medicine",
                       "Reproductive Endocrinology & Infertility",
                       "Urogynecology"), each = 3),
  year  = rep(c(2013, 2018, 2023), times = 4),
  count = c(1900, 2200, 2600,   # Gyn Onc
            1500, 1900, 2400,   # MFM
            1100, 1300, 1600,   # REI
             900, 1400, 2100)   # Urogyn
)
head(counts)
```

Wide data frames (a `subspecialty` column plus one column per year) and
matrices (subspecialty rows, year columns) are accepted too.

## 2. The denominator: total female population

The denominator is the **total female population** for each year. If you have a
Census API key you can fetch it directly (ACS table `B01001`, variable
`B01001_026E`):

```{r census, eval = FALSE}
# Requires CENSUS_API_KEY and an internet connection.
fem_pop <- mysterycall_census_female_population(
  years  = c(2013, 2018, 2023),
  survey = "acs5"            # 5-year: available for every year, incl. 2020
)
fem_pop
```

The ACS **1-year** table was not released for 2020; when you request
`survey = "acs1"`, `fill_2020` chooses whether to substitute the 5-year value
(default), return `NA`, or error.

For a reproducible vignette we just supply the values directly — a year-named
vector, an ordered vector, or a `data.frame(year, population)` all work:

```{r denominator}
fem_pop <- c(`2013` = 160477237, `2018` = 165553033, `2023` = 168731234)
```

> These illustrative denominators are placeholders — swap in your own cited
> Census figures.

## 3. The trend figure

```{r trend}
p <- mysterycall_subspecialist_trend(
  counts,
  population       = fem_pop,
  numerator_source = "Illustrative counts (vignette)",
  accessed         = "2026-08-02"
)
p
```

The returned object's `$data` carries every computed value, so nothing is
hidden behind the plot:

```{r trend-data}
head(p$data)
```

### Confidence intervals

A per-100,000 rate is a count over a population, so it has an exact Poisson
confidence interval. Set `conf_level` to draw it as a band and add
`density_low` / `density_high` to `$data`:

```{r trend-ci}
p_ci <- mysterycall_subspecialist_trend(counts, population = fem_pop,
                                        conf_level = 0.95)
head(p_ci$data[, c("subspecialty", "year", "density",
                   "density_low", "density_high")])
```

## 4. The two-point infographic

When you only need the endpoints, the infographic renders a badge per
subspecialty with the percent change derived from the two densities:

```{r infographic, fig.height = 3.4}
ends <- subset(counts, year %in% c(2013, 2023))
d13 <- ends$count[ends$year == 2013] / fem_pop["2013"] * 1e5
d23 <- ends$count[ends$year == 2023] / fem_pop["2023"] * 1e5

mysterycall_subspecialist_infographic(
  subspecialty = unique(ends$subspecialty),
  start = as.numeric(d13),
  end   = as.numeric(d23)
)
```

## 5. Provenance travels with the figure

Every figure records where its numbers came from. The record is attached to
the object and printed on the figure as a source caption:

```{r provenance}
attr(p, "provenance")
```

Saving a figure also writes provenance sidecars next to the image — a
human-readable `.provenance.txt` and, when **jsonlite** is installed, a
machine-readable `.provenance.json` (schema `mysterycall/provenance`) carrying
the record and the per-point table:

```{r save, eval = FALSE}
mysterycall_subspecialist_trend(
  counts, population = fem_pop,
  numerator_source = "ABOG certified-diplomate counts, 2013-2023",
  accessed = "2026-08-02",
  output_path = "subspecialists_per_100k.png"
)
#> Saved: subspecialists_per_100k.png
#> Provenance: subspecialists_per_100k.provenance.txt
#> Provenance: subspecialists_per_100k.provenance.json
```

That trio — image, human-readable provenance, machine-readable provenance —
is what makes the figure reproducible in an analysis pipeline.
