---
title: "Assembling supplementary digital content"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assembling supplementary digital content}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE,
  fig.width = 6.5, fig.height = 3.6, fig.align = "center"
)
sdc_dir <- file.path(tempdir(), "sdc")
dir.create(sdc_dir, showWarnings = FALSE)
```

Journals rarely print a mystery-caller study's full model tables, sensitivity
analyses, reporting checklist, and participant-flow diagram in the main text —
those go in the **supplementary digital content (SDC)**, uploaded as separate
files. This vignette assembles that package: the supplementary tables and
figures, a STROBE checklist, and the export steps, each generated from the
fitted models so the SDC stays consistent with the manuscript and is trivial to
regenerate when the data change.

```{r}
library(mysterycall)
```

## 1. The analysis behind the SDC

A mystery-caller study of insurance-based access has two outcomes: whether an
appointment was offered, and the wait in business days among offers. We simulate
a small paired call log and fit both models — a logistic GLMM for the offer and
a Poisson GLMM for the wait.

```{r fit}
set.seed(2026)
n <- 120
d <- data.frame(
  npi        = rep(sprintf("1%09d", seq_len(n)), each = 2),
  insurance  = rep(c("Commercial", "Medicaid"), n),
  region     = rep(sample(c("Northeast", "South", "West"), n, TRUE), each = 2)
)
med <- d$insurance == "Medicaid"
d$appt_offered <- rbinom(nrow(d), 1, plogis(1.1 - 0.9 * med))
d$wait_days    <- ifelse(d$appt_offered == 1,
                         rpois(nrow(d), ifelse(med, 18, 12)), NA)

offer_crude    <- mysterycall_logistic_model(d, "appt_offered", "insurance", "npi")
offer_adjusted <- mysterycall_logistic_model(
  d, "appt_offered", c("insurance", "region"), "npi")
wait_model <- mysterycall_poisson_model(
  d[d$appt_offered == 1, ], outcome = "wait_days",
  predictors = "insurance", random_intercept = "npi")
```

## 2. Supplementary tables

**Table S1 — model estimates (crude vs. adjusted).**
`mysterycall_multi_model_table()` places nested models side by side, the usual
"did adjustment move the estimate?" supplementary table.

```{r tableS1}
tS1 <- mysterycall_multi_model_table(
  list(Crude = offer_crude, Adjusted = offer_adjusted))
tS1
```

**Table S2 — the disparity, with confidence intervals.**

```{r tableS2}
tS2 <- mysterycall_disparities_table(d, "appt_offered", "insurance",
                                     ref_group = "Commercial")
tS2
```

**Table S3 — adjusted effect on the absolute scale.**
`mysterycall_combined_results_table()` reports the wait model's rate ratio
alongside the implied difference in days.

```{r tableS3}
tS3 <- mysterycall_combined_results_table(wait_model)
tS3
```

**Table S4 — this study in the context of prior work.**
`mysterycall_literature_table()` stacks your estimate against published ones.

```{r tableS4}
prior <- data.frame(
  author = c("Pollack 2016", "Sharma 2025"),
  year = c(2016, 2025),
  insurance_comparison = "Medicaid vs Commercial",
  n = c(1200, 480),
  or = c(0.42, 0.55), ci_lower = c(0.35, 0.40), ci_upper = c(0.51, 0.75),
  specialty = c("Urology", "OBGYN"))
mysterycall_literature_table(prior)
```

**A one-file bundle.** `mysterycall_supplemental_tables()` writes every model
table into a single formatted workbook for upload (needs the openxlsx package).

```{r bundle, eval = requireNamespace("openxlsx", quietly = TRUE)}
xlsx <- mysterycall_supplemental_tables(
  logistic_fit = offer_adjusted, poisson_fit = wait_model,
  file = file.path(sdc_dir, "supplemental_tables.xlsx"), overwrite = TRUE)
basename(xlsx)
```

## 3. Supplementary figures, at journal specification

**Figure S1 — forest plot of the offer model.**

```{r figS1}
figS1 <- mysterycall_forest_plot(offer_adjusted, x_label = "Odds ratio (95% CI)")
figS1
```

**Figure S2 — incidence-rate-ratio plot of the wait model.**

```{r figS2}
mysterycall_irr_plot(wait_model)
```

**Figure S3 — cumulative time to an appointment**, the correct single-contact
time-to-event display:

```{r figS3, fig.height = 3.4}
curve <- mysterycall_cumulative_access_curve(
  d, time_col = "wait_days", offered_col = "appt_offered",
  group_col = "insurance", horizon = 45, plot = TRUE)
curve$plot
```

**Save at the journal's column width.** `mysterycall_save_green_journal_figure()`
writes a figure at *Obstetrics & Gynecology* specifications (sized to a single or
double column, at print DPI) and returns the path.

```{r savefig}
paths <- mysterycall_save_green_journal_figure(
  figS1, file.path(sdc_dir, "figS1_forest"), layout = "single_column")
basename(paths)
```

## 4. The reporting checklist

Observational studies ship a STROBE checklist. `mysterycall_strobe_checklist()`
builds one from the fitted model, pre-filling the items it can verify:

```{r strobe}
head(mysterycall_strobe_checklist(wait_model), 8)
```

The participant-flow diagram (a near-universal Figure S1) comes from
`mysterycall_flowchart()` / `mysterycall_strobe_flow()`; it renders with the
DiagrammeR package:

```{r flow, eval = FALSE}
mysterycall_flowchart(
  counts = c("Sampled" = 240, "Reached" = 226, "Analyzed" = 218))
```

## 5. Export for upload

Each table can be written in whatever format the journal wants. A CSV per table
is the most portable:

```{r csv}
write.csv(as.data.frame(tS1), file.path(sdc_dir, "TableS1_models.csv"),
          row.names = FALSE)
write.csv(tS2, file.path(sdc_dir, "TableS2_disparities.csv"), row.names = FALSE)
```

For a formatted Word document (with a results paragraph on top), use
`mysterycall_export_results_docx()` — it needs the officer/flextable packages:

```{r docx, eval = FALSE}
mysterycall_export_results_docx(
  table = tS2, title = "Table S2. Insurance-based access disparities",
  output_path = file.path(sdc_dir, "TableS2.docx"))
```

## 6. A manifest

Finally, list what the SDC package contains — reviewers appreciate a manifest,
and it doubles as a build check.

```{r manifest}
files <- list.files(sdc_dir)
data.frame(
  file = files,
  kind = ifelse(grepl("\\.csv$", files), "table",
         ifelse(grepl("\\.(png|pdf|tiff)$", files), "figure", "bundle")))
```

## Recap

Every piece of a mystery-caller study's supplementary digital content — the
crude-vs-adjusted model table, the disparity table, the absolute-scale effect,
the literature comparison, the forest / IRR / cumulative-access figures at
journal specification, the STROBE checklist, and the export to workbook / Word /
CSV — comes from the same fitted models with one call apiece, so the SDC
regenerates itself whenever the analysis is re-run.
