---
title: "Assembling supplementary digital content"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assembling supplementary digital content}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", warning = FALSE, message = FALSE,
  fig.width = 6.5, fig.height = 3.6, fig.align = "center"
)
sdc_dir <- file.path(tempdir(), "sdc")
dir.create(sdc_dir, showWarnings = FALSE)
```

Journals rarely print a mystery-caller study's full model tables, sensitivity
analyses, reporting checklists, participant flow, and maps in the main text —
those go in the **supplementary digital content (SDC)**, uploaded as separate
files. This vignette assembles that package: model equations, supplementary
tables and figures, a missingness analysis, a per-caller evaluation, maps, and
two reporting checklists, each generated from the fitted models and the call log
so the SDC stays consistent with the manuscript and regenerates itself when the
data change.

```{r}
library(mysterycall)
```

## 1. The analysis behind the SDC

A mystery-caller study of insurance-based access has two outcomes — whether an
appointment was offered and the wait in business days among offers — and a call
log carrying the caller, the call date, and the practice location. We simulate
it, then fit a logistic GLMM for the offer and a Poisson GLMM for the wait.

```{r fit}
set.seed(2026)
n <- 150
cities <- list(CO = "Denver", TX = "Dallas", NY = "New York",
               CA = "Los Angeles", FL = "Miami", IL = "Chicago")
st <- sample(names(cities), n, TRUE)
d <- data.frame(
  npi        = rep(sprintf("1%09d", seq_len(n)), each = 2),
  insurance  = rep(c("Commercial", "Medicaid"), n),
  caller     = rep(sample(paste0("RA", 1:5), n, TRUE), each = 2),
  call_date  = rep(as.Date("2026-01-05") + sample(0:40, n, TRUE), each = 2),
  state      = rep(st, each = 2),
  city       = rep(unlist(cities)[st], each = 2),
  area       = rep(sample(c("Metro", "Nonmetro"), n, TRUE), each = 2),
  degree     = rep(sample(c("MD", "DO"), n, TRUE, c(.8, .2)), each = 2)
)
med <- d$insurance == "Medicaid"
d$reached <- rbinom(nrow(d), 1, 0.92) == 1   # some calls never reached a live office
d$appt_offered <- rbinom(nrow(d), 1, plogis(1.1 - 0.9 * med + 0.3 * (d$area == "Metro")))
# a few calls never resolved the wait (incomplete) -> missing outcome
d$wait_days <- ifelse(d$appt_offered == 1 & rbinom(nrow(d), 1, 0.9) == 1,
                      rpois(nrow(d), ifelse(med, 18, 12)), NA)

offer_crude    <- mysterycall_logistic_model(d, "appt_offered", "insurance", "npi")
offer_adjusted <- mysterycall_logistic_model(d, "appt_offered",
                                             c("insurance", "area"), "npi")
wait_model <- mysterycall_poisson_model(
  d[d$appt_offered == 1 & !is.na(d$wait_days), ], outcome = "wait_days",
  predictors = "insurance", random_intercept = "npi")
```

## 2. Model equations

Supplementary methods print the statistical model. `mysterycall_model_equation()`
renders a fitted GLM(M) as LaTeX — no `equatiomatic` dependency — reading the
link, terms, and random intercept off the fit.

```{r eq, results = "asis"}
cat(mysterycall_model_equation(offer_adjusted), "\n\n")
cat(mysterycall_model_equation(wait_model))
```

Pass `coefficients = TRUE` for the fitted form with numbers in place of the
$\beta$s; the symbolic form carries a `legend` attribute mapping each
coefficient to its term.

## 3. Supplementary tables

**Table S1 — model estimates (crude vs. adjusted).**

```{r tableS1}
tS1 <- mysterycall_multi_model_table(
  list(Crude = offer_crude, Adjusted = offer_adjusted))
tS1
```

**Table S2 — the disparity, with confidence intervals.**

```{r tableS2}
tS2 <- mysterycall_disparities_table(d, "appt_offered", "insurance",
                                     ref_group = "Commercial")
tS2
```

**Table S3 — adjusted effect on the absolute scale.**

```{r tableS3}
mysterycall_combined_results_table(wait_model)
```

**Table S4 — this study against prior work.**

```{r tableS4}
prior <- data.frame(
  author = c("Pollack 2016", "Sharma 2025"), year = c(2016, 2025),
  insurance_comparison = "Medicaid vs Commercial", n = c(1200, 480),
  or = c(0.42, 0.55), ci_lower = c(0.35, 0.40), ci_upper = c(0.51, 0.75),
  specialty = c("Urology", "OBGYN"))
mysterycall_literature_table(prior)
```

## 4. Missing-data analysis

Reviewers want to know whether missingness is benign.
`mysterycall_missing_data_analysis()` quantifies it and tests whether it depends
on the exposure — the difference between "missing completely at random" and a
threat to validity.

```{r missing}
miss <- mysterycall_missing_data_analysis(
  d, outcome_col = "wait_days", group_col = "insurance")
miss$summary
```

## 5. Per-caller evaluation

Standardization is central to a covert audit, so the SDC should show each caller
behaved comparably. `mysterycall_call_productivity()` summarizes volume and
outcomes by caller; large between-caller swings in the acceptance rate would flag
a standardization problem.

```{r caller}
mysterycall_call_productivity(
  d, caller_col = "caller", date_col = "call_date", outcome_col = "appt_offered")
```

## 6. Supplementary figures, at journal specification

**Figure S1 — forest plot of the offer model.**

```{r figS1}
figS1 <- mysterycall_forest_plot(offer_adjusted, x_label = "Odds ratio (95% CI)")
figS1
```

**Figure S2 — cumulative time to an appointment**, the correct single-contact
time-to-event display:

```{r figS2, fig.height = 3.4}
mysterycall_cumulative_access_curve(
  d, time_col = "wait_days", offered_col = "appt_offered",
  group_col = "insurance", horizon = 45, plot = TRUE)$plot
```

`mysterycall_save_green_journal_figure()` writes any figure at *Obstetrics &
Gynecology* column width and print DPI:

```{r savefig}
basename(mysterycall_save_green_journal_figure(
  figS1, file.path(sdc_dir, "figS1_forest"), layout = "single_column"))
```

## 7. Maps

**An interactive practice map.** Geocode the roster and drop the practices onto a
Leaflet map — an interactive supplement that print cannot show.

```{r leaflet, eval = requireNamespace("leaflet", quietly = TRUE), fig.height = 4}
practices <- unique(d[, c("npi", "city", "state", "insurance")])
g <- mysterycall_geocode_city_state(practices$city, practices$state)
practices$lat <- g$lat; practices$lon <- g$lon
leaflet::addCircleMarkers(
  leaflet::addTiles(leaflet::leaflet(practices)),
  lng = ~lon, lat = ~lat, radius = 4, stroke = FALSE, fillOpacity = 0.6,
  popup = ~paste0(city, ", ", state))
```

**A choropleth by region/district.** Assign every state its ACOG district with
`mysterycall_assign_region()` and shade a US map:

```{r regionmap, eval = requireNamespace("maps", quietly = TRUE), fig.height = 4}
library(ggplot2)
us <- map_data("state")
districts <- data.frame(region = tolower(state.name),
                        district = mysterycall_assign_region(state.abb))
choro <- merge(us, districts, by = "region")
choro <- choro[order(choro$order), ]
ggplot(choro, aes(long, lat, group = group, fill = district)) +
  geom_polygon(colour = "white", linewidth = 0.2) +
  coord_quickmap() + theme_void() +
  labs(fill = "ACOG district", title = "Sampling frame by ACOG district")
```

(`mysterycall_region_labels()` returns the same state-to-region assignment with
centroids for labelling.)

## 8. Reporting checklists

Observational studies ship a **STROBE** checklist, built from the fitted model:

```{r strobe}
head(mysterycall_strobe_checklist(wait_model), 6)
```

But mystery-caller studies are *simulated-patient* research, and STROBE omits the
covert-methodology items reviewers ask for — why the covert method was justified,
how callers were trained, how detection was handled, how the deception was
ethically managed. `mysterycall_crisp_checklist()` supplies that companion:

```{r crisp}
crisp <- mysterycall_crisp_checklist()
crisp[crisp$section %in% c("Callers", "Detection", "Ethics"),
      c("section", "item")]
```

The participant-flow diagram (Figure S-flow) comes from
`mysterycall_flowchart()`, which renders with the DiagrammeR package:

```{r flow, eval = FALSE}
mysterycall_flowchart(
  counts = c("Sampled" = 240, "Reached" = 226, "Analyzed" = 218))
```

## 9. Sensitivity analyses

A covert audit invites a predictable set of "but what if..." questions, and the
SDC is where you answer them. Two are near-universal.

**Non-response bounds.** Not every call reaches a live office, so the offer rate
is not point-identified. Rather than quietly condition on completed calls,
`mysterycall_outcome_bounds()` reports the assumption-free interval — assign every
unreached call first to failure, then to success:

```{r bounds}
mysterycall_outcome_bounds(d, "appt_offered", observed = "reached")
```

**Leave-one-caller-out.** Confirm the exposure effect does not hinge on a single
research assistant by refitting with each caller dropped in turn:

```{r loco}
fit_glm <- glm(appt_offered ~ insurance + area, binomial, d)
mysterycall_leave_one_out(fit_glm, d, group = "caller", term = "insuranceMedicaid")
```

Other analyses worth including, each a single call: the wait model under an
alternative count family (`mysterycall_compare_count_families()`), calendar vs.
business days (`mysterycall_calendar_sensitivity()`), a stricter outcome
definition (re-fit on "offer with the sampled physician"), and
Benjamini–Hochberg-adjusted p-values
(`mysterycall_multiple_comparison_adjust()`).

## 10. Baseline characteristics (Table 1)

Reviewers want to know who was sampled. `mysterycall_table1()` builds a
publication-ready baseline table at the practice level:

```{r table1}
practices <- unique(d[, c("npi", "state", "area", "degree")])
mysterycall_table1(practices, covariates = c("state", "degree"),
                   stratify_by = "area", include_overall = TRUE)
```

## 11. Full model output: ICC and variance components

The SDC should report more than the exposure coefficient — the clustering matters.
`mysterycall_icc()` gives the intraclass correlation (how much of the variation
sits between practices), and `mysterycall_random_effect_variance()` the variance
components:

```{r icc}
mysterycall_icc(wait_model)
mysterycall_random_effect_variance(wait_model$model)
```

## 12. A priori power

Journals require the pre-specified power calculation.
`mysterycall_poisson_power()` sizes the wait-time comparison for a target
incidence-rate ratio:

```{r power}
mysterycall_poisson_power(irr = 1.40, lambda_ref = 14, power = 0.80,
                          both_arms = TRUE)
```

## 13. Model diagnostics

A residual figure justifies the model.
`mysterycall_plot_residuals()` draws the diagnostic panel, and the fitted object
already carries the overdispersion check:

```{r diagnostics, fig.height = 3.2}
mysterycall_plot_residuals(wait_model)
```

## 14. Reproducibility

Finally, a reproducibility appendix — the R and package versions the analysis ran
under, so the SDC is a complete, re-runnable record:

```{r repro}
mysterycall_session_snapshot(
  file = file.path(sdc_dir, "session_snapshot.txt"), quiet = TRUE,
  notes = "Supplementary digital content build")
```

## 15. Export and manifest

Write each artifact in the format the journal wants — a CSV per table is the most
portable, `mysterycall_export_results_docx()` gives formatted Word (needs
officer/flextable), and `mysterycall_supplemental_tables()` bundles every model
table into one workbook (needs openxlsx):

```{r export}
write.csv(as.data.frame(tS1), file.path(sdc_dir, "TableS1_models.csv"),
          row.names = FALSE)
write.csv(tS2, file.path(sdc_dir, "TableS2_disparities.csv"), row.names = FALSE)
write.csv(as.data.frame(crisp), file.path(sdc_dir, "ChecklistS1_CRiSP.csv"),
          row.names = FALSE)
```

```{r bundle, eval = requireNamespace("openxlsx", quietly = TRUE)}
mysterycall_supplemental_tables(
  logistic_fit = offer_adjusted, poisson_fit = wait_model,
  file = file.path(sdc_dir, "supplemental_tables.xlsx"), overwrite = TRUE)
```

Finally, a manifest — reviewers appreciate it, and it doubles as a build check:

```{r manifest}
files <- list.files(sdc_dir)
data.frame(
  file = files,
  kind = ifelse(grepl("\\.csv$", files), "table",
         ifelse(grepl("\\.(png|pdf|tiff)$", files), "figure",
         ifelse(grepl("\\.xlsx$", files), "bundle", "other"))))
```

## Recap

Every piece of a mystery-caller study's supplementary digital content — the model
equations, a baseline Table 1, the crude-vs-adjusted and disparity tables, the
ICC and variance components, the missingness analysis, the per-caller evaluation,
the a-priori power calculation, the forest, cumulative-access, and residual
figures at journal specification, the practice and district maps, the STROBE
**and** CRiSP reporting checklists, non-response and leave-one-caller-out
sensitivity analyses, and a reproducibility appendix — comes from the same fitted
models and call log with one call apiece, so the SDC regenerates itself whenever
the analysis is re-run.
