---
title: "The Two-Part Model and the Denominator Trap"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    df_print: kable
description: >
  Access has two stages, whether an appointment is offered and how long the wait
  is among offers. Each stage needs its own denominator, and exclusion criteria
  that absorb the no-offer outcome will silently destroy the first stage.
vignette: >
  %\VignetteIndexEntry{The Two-Part Model and the Denominator Trap}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE, purl=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = eval_vignettes
)
library(mysterycall)
```

## Access is two questions, not one

"How long is the wait?" is only half of an access study. An office that never
offers an appointment has a wait of neither zero nor infinity; it has no wait at
all, and folding it in either direction misstates the finding.

The two-part (hurdle) structure keeps the questions separate:

1. **Offer stage.** Was an appointment offered? Binary, over everyone who was
   called.
2. **Wait stage.** Among appointments offered, how long? Count, over offers
   only.

Reported together they answer the question a patient actually has. Reported as a
single mean wait, they can move in opposite directions and cancel: an arm that
refuses more patients can post a *shorter* mean wait, because the refusals never
enter the average.

## The trap

Here is the part that is easy to miss and expensive to get wrong.

Mystery-caller protocols carry an exclusion vocabulary: wrong number, voicemail,
closed system, retired. These are legitimate. The trap is that such vocabularies
often quietly include categories that are **outcomes rather than exclusions** --
"not accepting new patients", "requires a referral first", "would not schedule
without records".

Those are refusals. If the protocol codes them as exclusions, they are removed
from the cohort, and the offer stage loses almost all of its variation.

```{r trap}
calls <- data.frame(
  call_id = sprintf("C%02d", 1:20),
  reason  = c(rep("Able to contact", 12),
              rep("Not accepting new patients", 4),   # outcome, coded as exclusion
              rep("Wrong number listed", 2),          # genuine exclusion
              rep("Went to voicemail", 2)),           # genuine exclusion
  stringsAsFactors = FALSE
)
calls$offered <- calls$reason == "Able to contact"

protocol_eligible <- calls$reason == "Able to contact"
all_called        <- rep(TRUE, nrow(calls))

data.frame(
  denominator = c("source (everyone called)", "protocol-eligible only"),
  n           = c(sum(all_called), sum(protocol_eligible)),
  offered     = c(sum(calls$offered[all_called]), sum(calls$offered[protocol_eligible])),
  offer_rate  = c(round(100 * mean(calls$offered[all_called]), 1),
                  round(100 * mean(calls$offered[protocol_eligible]), 1))
)
```

On the eligible cohort the offer rate is 100 percent. There is nothing left to
model: every office that survived the exclusion criteria offered an appointment,
because *not offering one was itself an exclusion criterion*. Fit a logistic
model there and it will fail to converge, or converge on nothing, and the
natural reading is "no disparity in access" when the truth is that the question
was never asked.

The four "not accepting new patients" calls are the finding. They belong in the
denominator.

## Choosing each denominator deliberately

```{r denominators}
# Stage 1: offer. Everyone called who could in principle have been offered
# something -- genuine unreachables (wrong number, voicemail) removed, refusals
# retained, because a refusal is an outcome.
genuine_exclusion <- calls$reason %in% c("Wrong number listed", "Went to voicemail")
offer_denominator <- !genuine_exclusion

# Stage 2: wait. Offers only.
wait_denominator <- calls$offered

data.frame(
  stage = c("offer", "wait"),
  denominator_n = c(sum(offer_denominator), sum(wait_denominator)),
  definition = c("called, minus unreachable-by-construction",
                 "appointments offered")
)
```

Stated as a rule: **an exclusion removes calls where the question could not be
asked; an outcome records the answer when it was.** Wrong number means the
office was never reached, so no access question was posed. "Not accepting new
patients" means the office was reached and said no.

## Reporting both stages

```{r report}
offer_rate <- mean(calls$offered[offer_denominator])
sprintf("Offer stage: %d of %d calls (%.0f%%) yielded an appointment.",
        sum(calls$offered[offer_denominator]), sum(offer_denominator), 100 * offer_rate)
```

A wait-time result quoted without its offer rate is not interpretable, because
the reader cannot tell whether a short wait reflects good access or heavy
attrition upstream. Report the pair, always, and report the denominator of each.

`mysterycall_hurdle_wait()` fits both stages together and keeps the denominators
distinct; `mysterycall_twopart_power()` does power for the same structure, which
matters because the two stages have very different effective sample sizes.

## Diagnosing the trap in an existing study

Three checks, cheap to run, that catch this before modelling:

```{r diagnose}
diagnose <- function(reason, offered, eligible) {
  data.frame(
    check = c("offer-stage variation within eligible cohort",
              "exclusion categories that read as outcomes",
              "share of calls removed before the offer stage"),
    value = c(
      sprintf("%.0f%% offered", 100 * mean(offered[eligible])),
      paste(intersect(unique(reason),
                      c("Not accepting new patients", "Requires referral",
                        "Would not schedule without records")), collapse = "; "),
      sprintf("%.0f%%", 100 * mean(!eligible))
    )
  )
}
diagnose(calls$reason, calls$offered, protocol_eligible)
```

If the first row is at or near 100 percent, the offer stage has been absorbed by
the exclusion criteria and the denominator needs revisiting before any model is
fit. If the second row is non-empty, you have found where.

## Where this fits

- `mysterycall_hurdle_wait()` for the joint model, and the `logistic-model` and
  `linear-mixed-models` vignettes for each stage on its own.
- The `cohort-provenance` vignette for keeping the source, eligible and analytic
  populations as separate objects, which is what makes the offer denominator
  recoverable at all.
- `mysterycall_access_cascade()` for reporting the stages as a cascade when there
  are more than two.
