---
title: "End-to-End Mystery-Caller Workflow Orchestration"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    df_print: kable
description: >
  Run the complete mystery-caller pipeline in one call with
  mysterycall_run_workflow(), or step through Phase 1 cleaning,
  caller-list splitting, Phase 2 cleaning, quality-check export, and
  coverage monitoring individually.
vignette: >
  %\VignetteIndexEntry{End-to-End Mystery-Caller Workflow Orchestration}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
eval_vignettes <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = eval_vignettes)
library(mysterycall)
```

```{r load-pkg, eval=FALSE}
library(mysterycall)
```

---

# 1. Workflow Overview

Mystery-caller studies are a form of audit research in which trained research
assistants — referred to here as *field callers* or *lab assistants* — pose as
prospective patients and attempt to schedule appointments with physicians
identified from a national provider database. The `mysterycall` package
manages every computational step of that pipeline, from pulling National
Provider Identifier (NPI) records through writing the final analysis-ready
data set.

## 1.1 The full pipeline as a directed acyclic graph

```
  ┌─────────────────────┐
  │   NPI Search        │  mysterycall_search_and_process_npi() / NPPES API
  │  (taxonomy terms)   │
  └──────────┬──────────┘
             │  raw provider roster
             ▼
  ┌─────────────────────┐
  │  Phase 1 Cleaning   │  mysterycall_clean_phase1()
  │  · dedup names      │
  │  · validate phones  │
  │  · flag missings    │
  │  · generate IDs     │
  └──────────┬──────────┘
             │  clean roster + audit JSON
             ▼
  ┌─────────────────────┐
  │  NPI Validation     │  mysterycall_validate_npi()
  │  (Luhn checksum)    │  mysterycall_check_duplicates()
  └──────────┬──────────┘
             │  validated roster
             ▼
  ┌─────────────────────┐
  │  Split into         │  mysterycall_split_and_save()
  │  Caller Workbooks   │
  └──────┬──────────────┘
         │  one Excel file per caller
         ▼
  ┌─────────────────────┐
  │  [Field Callers]    │  (human step — callers make calls, fill Excel)
  │  make calls,        │
  │  record outcomes    │
  └──────────┬──────────┘
             │  returned Excel workbooks
             ▼
  ┌─────────────────────┐
  │  Phase 2 Cleaning   │  mysterycall_rename_columns()
  │  · harmonise cols   │  mysterycall_clean_phase2()
  │  · parse outcomes   │
  └──────────┬──────────┘
             │  standardised outcome data
             ▼
  ┌─────────────────────┐
  │  Quality Check      │  mysterycall_save_quality_table()
  │  per-caller stats   │
  └──────────┬──────────┘
             │  QA snapshot
             ▼
  ┌─────────────────────┐
  │  Coverage Monitor   │  mysterycall_not_contacted_states()
  │  · which states     │
  │    have zero calls? │
  └─────────────────────┘
```

## 1.2 Phase 1 vs Phase 2: what each phase does

**Phase 1** happens *before* callers pick up the phone. Its job is to turn
a raw export from NPPES (or a similar roster) into a clean, validated list
that is ready to be handed to field callers. Concretely, Phase 1:

* Strips whitespace and normalises letter case in provider names.
* Validates and formats phone numbers.
* Assigns a reproducible random study ID to every row, including rows where
  the NPI is missing.
* Flags data-quality problems (empty names, missing last names, duplicate
  entries) without silently dropping any rows.
* Optionally duplicates each row so that a single provider appears once for
  Medicaid and once for private insurance — a common design in disparity
  studies.
* Writes a machine-readable audit JSON to disk so that every decision is
  traceable.

**Phase 2** happens *after* callers return their completed workbooks. Its job
is to harmonise the heterogeneous column names that inevitably accumulate
across five or ten individual Excel files and to normalise the free-text
outcome values (e.g., "yes", "YES", "Y", "offered" → a single canonical
code) so that the merged file is analysis-ready.

## 1.3 Two ways to run the pipeline

**Option A — One-call orchestration (recommended for routine studies):**

```{r one-call, eval=FALSE}
result <- mysterycall_run_workflow(
  taxonomy_terms      = c("207V00000X", "207VB0002X"),
  name_data           = provider_names_df,
  phase1_data         = raw_roster,
  lab_assistant_names = c("Alice", "Bob", "Carol", "Dana", "Eve"),
  output_directory    = "/network/study/phase1_output",
  phase2_data         = "/network/study/returned_workbooks",
  phase2_output_directory = "/network/study/phase2_output",
  quality_check_path  = "/network/study/qa/quality_table.csv",
  all_states          = state.abb,
  verbose             = TRUE
)
```

**Option B — Step-by-step (recommended when debugging or adapting the
workflow to an unusual study design):**

Each function documented in Sections 2–7 below can be called independently.
Results flow from one function to the next as ordinary R data frames, so you
can inspect, filter, or modify the intermediate outputs at any point.

---

# 2. Phase 1 Cleaning: `mysterycall_clean_phase1()`

Phase 1 cleaning is the most complex and consequential step in the pipeline.
Every row that passes through this function is tagged with a set of
`processing_flag_*` columns that record exactly why it was modified. This
makes the cleaning process auditable by anyone — a collaborator, a journal
reviewer, or future-you two years from now.

## 2.1 Required input: what `phase1_data` must contain

`phase1_data` must be a data frame (or a path to a CSV / Excel file) with at
least the following columns:

| Column | Description |
|---|---|
| `npi` | 10-digit NPI as a character string (may be `NA`) |
| `provider_first_name` | Provider's first name (may be empty) |
| `provider_last_name_legal_name` | Provider's legal last name |
| `provider_business_practice_location_address_telephone_number` | Phone number (raw format, may have dashes or dots) |

Additional columns (specialty, address, etc.) are passed through untouched.

## 2.2 Key parameters

```{r phase1-params, eval=FALSE}
result <- mysterycall_clean_phase1(
  phase1_data      = raw_roster,        # required: raw data frame or file path
  output_directory = "/study/output",   # required: directory for CSV + audit JSON
  output_format    = "csv",             # "csv" (default), "rds", or "parquet"
  verbose          = TRUE,              # print step-by-step progress
  duplicate_rows   = TRUE,             # TRUE = one row per insurance scenario
  id_seed          = 20240101L,         # integer seed for reproducible ID generation
  parent_cohort_hash = NULL             # link to a previous run's hash (optional)
)
```

**`duplicate_rows = TRUE`** is the correct setting for disparity studies
that test both Medicaid and private-insurance callers at the same practices.
When `TRUE`, every row in `phase1_data` is copied: one copy gets
`insurance = "Medicaid"` and the other gets `insurance = "Private"`. The
`processing_flag_is_duplicate` column records which rows are the duplicates.

**`id_seed`** controls the random number generator used to assign study IDs
to rows where the NPI is `NA`. Fixing this seed ensures that if you re-run
Phase 1 on the same data you get identical IDs — crucial for linking
records across pipeline runs.

**`parent_cohort_hash`** lets you chain audit trails. If Phase 1 was already
run for a pilot sample and you are now adding providers, pass the pilot's
`cohort_hash` here so the new audit JSON references the parent batch.

## 2.3 Columns added by Phase 1

Phase 1 adds the following `processing_flag_*` columns to the returned data
frame. These columns are `logical` (`TRUE`/`FALSE`).

| Column | Meaning when `TRUE` |
|---|---|
| `processing_flag_empty_name` | Both first and last name fields are blank or `NA` |
| `processing_flag_no_last_name` | Last name field is blank or `NA` (first name may still be present) |
| `processing_flag_generated_id` | NPI was missing; a random study ID was assigned by the package |
| `processing_flag_is_duplicate` | Row was created by `duplicate_rows = TRUE` and is a copy of another row |

These flags are never used to silently drop rows. Every flagged row is
retained in the output. The flags exist so that downstream analysts can
filter, weight, or report on data-quality issues without re-running the
cleaning step.

```{r phase1-flag-example, eval=FALSE}
# How many rows were assigned a generated ID?
sum(result$processing_flag_generated_id, na.rm = TRUE)

# Which rows have no last name?
result[result$processing_flag_no_last_name, c("provider_first_name",
                                               "provider_last_name_legal_name",
                                               "npi")]
```

## 2.4 The audit trail JSON

Every call to `mysterycall_clean_phase1()` writes a JSON file to
`output_directory`. The filename contains the cohort hash so it is unique
across runs. The JSON contains:

```json
{
  "cohort_hash": "a3f8c1d2...",
  "parent_cohort_hash": null,
  "timestamp": "2024-03-15T09:22:41Z",
  "input_rows": 847,
  "output_rows": 1694,
  "empty_names_count": 12,
  "rows_retained_pct": 100.0,
  "quality_metrics": {
    "completeness_npi": 0.961,
    "completeness_phone": 0.983,
    "completeness_names": 0.986
  },
  "parameters": {
    "duplicate_rows": true,
    "id_seed": 20240101,
    "output_format": "csv"
  }
}
```

Key fields:

* **`input_rows`** — number of rows in `phase1_data` before any processing.
* **`output_rows`** — number of rows in the returned data frame. If
  `duplicate_rows = TRUE` this will be approximately twice `input_rows`.
* **`empty_names_count`** — number of rows where both name fields were blank.
* **`rows_retained_pct`** — always 100 in standard runs; any value below 100
  signals that rows were filtered (which only happens if you set
  `drop_empty_names = TRUE`, a non-default option).
* **`quality_metrics`** — proportion of rows with non-missing values for each
  critical field. Use these to decide whether the roster is ready for field
  work.

## 2.5 Accessing the returned attributes

The returned data frame carries two attributes that link it to the audit
trail:

```{r phase1-attrs, eval=FALSE}
# The cohort hash uniquely identifies this batch
cohort_id <- attr(result, "cohort_hash")
cat("Cohort hash:", cohort_id, "\n")

# The path to the JSON file on disk
audit_path <- attr(result, "audit_trail_path")
cat("Audit trail written to:", audit_path, "\n")

# Read the audit trail back for inspection
audit <- jsonlite::read_json(audit_path, simplifyVector = TRUE)
cat("Completeness — NPI:  ", audit$quality_metrics$completeness_npi, "\n")
cat("Completeness — Phone:", audit$quality_metrics$completeness_phone, "\n")
```

Store `cohort_id` in your lab notebook or electronic data-capture system. It
is the stable reference that links every downstream output (split workbooks,
Phase 2 files, model results) back to this exact version of the cleaned roster.

## 2.6 Common mistake: missing required columns

If `phase1_data` does not contain an `npi` column, `mysterycall_clean_phase1()`
throws an informative error immediately:

```{r phase1-error, eval=FALSE}
# This will fail with a clear error message
bad_data <- data.frame(
  provider_first_name = "Jane",
  provider_last_name_legal_name = "Smith",
  phone = "303-555-0100"
  # 'npi' column is absent
)

mysterycall_clean_phase1(
  phase1_data      = bad_data,
  output_directory = tempdir()
)
#> Error in mysterycall_clean_phase1():
#>   Required column 'npi' not found in phase1_data.
#>   Found columns: provider_first_name, provider_last_name_legal_name, phone
#>   Please rename or add the 'npi' column before calling this function.
```

---

# 3. NPI Validation: `mysterycall_validate_npi()`

The National Provider Identifier is a 10-digit number assigned by CMS. The
last digit is a **Luhn checksum digit** — a single-digit error-detection code
that catches single-digit transcription errors and most transposition errors
(e.g., swapping adjacent digits). `mysterycall_validate_npi()` applies the
Luhn algorithm to every NPI in the data and marks invalid ones.

## 3.1 Why Luhn validation matters

A roster built from web scraping, OCR, or manual data entry will contain
transcription errors. An invalid NPI cannot be verified against the NPPES
database, meaning you may end up calling a provider who does not exist or,
worse, a provider with a different specialty than you intended to sample. A
few minutes of Luhn validation before the field work starts prevents these
problems entirely.

## 3.2 Using the function

```{r npi-validate, eval=FALSE}
validated <- mysterycall_validate_npi(
  data       = phase1_result,
  npi_column = "npi"    # default; change if your column has a different name
)

# The function adds a logical column 'npi_valid'
table(validated$npi_valid, useNA = "ifany")
#>
#>  FALSE   TRUE   <NA>
#>     14   1678      2
```

Valid vs invalid — a quick illustration:

| NPI | Valid? | Reason |
|---|---|---|
| `1234567893` | TRUE | Passes Luhn checksum |
| `1234567890` | FALSE | Fails Luhn checksum |
| `123456789` | FALSE | Only 9 digits (must be exactly 10) |
| `12345678AB` | FALSE | Contains non-numeric characters |
| `NA` | NA | Missing — handled separately |

The function **always returns NPI as a character string**, even if the input
column contains integers. This is important because leading zeros (rare but
possible) would be silently dropped by numeric storage.

## 3.3 Combining with duplicate detection

After validation, use `mysterycall_check_duplicates()` to identify providers
who appear multiple times in the roster. Duplicate entries inflate your
apparent sample size and, if both entries are assigned to different callers,
result in the same provider being called twice — a protocol violation in most
IRB applications.

```{r npi-dups, eval=FALSE}
dup_report <- mysterycall_check_duplicates(
  data       = validated,
  id_column  = "npi",
  name_cols  = c("provider_first_name", "provider_last_name_legal_name")
)

# Rows where npi_count > 1 are duplicate providers
dups <- dup_report[dup_report$npi_count > 1, ]
cat(nrow(dups), "rows are duplicate NPI entries\n")

# Remove duplicates before splitting
clean_roster <- validated[!duplicated(validated$npi) | is.na(validated$npi), ]
```

---

# 4. Splitting into Caller Workbooks: `mysterycall_split_and_save()`

Once the roster is clean and validated, it must be divided among the field
callers. `mysterycall_split_and_save()` does this in a reproducible,
auditable way and writes one Excel workbook per caller.

## 4.1 Purpose and motivation

Imagine you have 500 providers and 5 lab assistants. You could manually
copy-paste rows into five spreadsheets, but that process is error-prone
(rows get skipped or duplicated), is not reproducible (results differ if
you redo it), and leaves no record of who was assigned what. This function
solves all three problems.

## 4.2 Key parameters

```{r split-params, eval=FALSE}
workbooks <- mysterycall_split_and_save(
  data                = clean_roster,
  lab_assistant_names = c("Alice", "Bob", "Carol", "Dana", "Eve"),
  output_directory    = "/study/workbooks",
  split_insurance_order = c("Medicaid", "Private"),
  seed                = 1978L       # default; controls assignment shuffle
)
```

**`lab_assistant_names`** — a character vector of caller names. Each name
becomes a separate Excel file: `Alice.xlsx`, `Bob.xlsx`, etc. The length of
this vector determines how many workbooks are produced.

**`split_insurance_order`** — controls the order in which insurance scenarios
appear within each workbook. If a provider appears once for Medicaid and once
for Private insurance (because `duplicate_rows = TRUE` was used in Phase 1),
the caller will see the Medicaid row first and the Private row second within
their workbook. This is useful for standardising the call script order.

**`seed = 1978`** — the default seed has been the package default since
version 1.0 and is the recommended value for all studies unless your protocol
specifies otherwise. Changing the seed changes which providers are assigned
to which caller; document any non-default seed in your methods section.

## 4.3 How assignment works: round-robin

Providers are assigned to callers in a **round-robin** pattern:

```
Provider 1 → Alice
Provider 2 → Bob
Provider 3 → Carol
Provider 4 → Dana
Provider 5 → Eve
Provider 6 → Alice   (wraps around)
Provider 7 → Bob
...
```

This means workloads are balanced within ±1 provider across all callers for
any roster size. The assignment is deterministic given the seed, the order of
rows in `data`, and the order of names in `lab_assistant_names`.

```{r workload-bar, echo=TRUE, eval=TRUE, fig.width=5, fig.height=2.8, fig.cap="Example round-robin workload distribution across five callers for a 198-provider roster. Differences of at most 1 provider guarantee balanced calling burden."}
callers   <- c("Alice", "Bob", "Carol", "Dana", "Eve")
n_total   <- 198L
base_load <- n_total %/% length(callers)
extra     <- n_total %% length(callers)
loads     <- c(rep(base_load + 1L, extra), rep(base_load, length(callers) - extra))

wb_df <- data.frame(
  caller = factor(callers, levels = callers),
  n      = loads
)

ggplot2::ggplot(wb_df, ggplot2::aes(x = caller, y = n, fill = caller)) +
  ggplot2::geom_col(width = 0.6) +
  ggplot2::geom_text(ggplot2::aes(label = n), vjust = -0.4, size = 3.5,
                      fontface = "bold") +
  ggplot2::scale_fill_brewer(palette = "Set2", guide = "none") +
  ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15)),
                               limits = c(0, NA)) +
  ggplot2::labs(x = NULL, y = "Providers assigned") +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(panel.grid.major.x = ggplot2::element_blank())
```

## 4.4 Verifying the split: no omissions, no duplicates

Before sending workbooks to callers, verify that every provider appears in
exactly one workbook:

```{r split-verify, eval=FALSE}
# Read all workbooks back and combine
all_workbooks <- lapply(workbooks$file_paths, readxl::read_excel)
combined      <- dplyr::bind_rows(all_workbooks)

# Check 1: same number of rows as the input
stopifnot(nrow(combined) == nrow(clean_roster))

# Check 2: no NPI appears in more than one workbook
npi_counts <- table(combined$npi)
if (any(npi_counts > 1)) {
  warning("Some NPIs appear in more than one workbook!")
  print(npi_counts[npi_counts > 1])
} else {
  message("Split verified: every provider appears in exactly one workbook.")
}
```

---

# 5. Phase 2 Cleaning: `mysterycall_clean_phase2()` and `mysterycall_rename_columns()`

After field callers return their completed workbooks, the returned files
must be merged and standardised. This is Phase 2. The core challenge is that
callers — despite receiving identical templates — often rename columns,
insert their own columns, or spell outcome values differently. Phase 2
handles all of this programmatically.

## 5.1 Why column names are always inconsistent

Even with a rigid Excel template, callers rename columns in practice:

| Caller | Column they return | Intended column |
|---|---|---|
| Alice | `Appointment Offered?` | `appointment_offered` |
| Bob | `Appt_Offered` | `appointment_offered` |
| Carol | `appointment offered` | `appointment_offered` |
| Dana | `APPT OFFERED` | `appointment_offered` |
| Eve | `appointment_offered` | `appointment_offered` ✓ |

`mysterycall_rename_columns()` resolves this via **fuzzy substring matching**.
You provide a list of target substrings to look for and the standardised
names to replace them with.

## 5.2 `mysterycall_rename_columns()`: fuzzy column name matching

```{r rename-cols, eval=FALSE}
# Standardise column names across all returned workbooks
renamed <- mysterycall_rename_columns(
  data           = raw_returned_workbook,
  target_strings = c("appt", "wait", "insurance", "outcome", "caller"),
  new_names      = c("appointment_offered", "wait_days",
                     "insurance_type", "call_outcome", "caller_name")
)
```

**`target_strings`** — a character vector of substrings (case-insensitive)
to search for in the column names of `data`. The function looks for each
substring anywhere in the column name.

**`new_names`** — a character vector of the same length as `target_strings`.
Each element is the standardised name to assign when the corresponding
substring is found.

If a target substring does not match any column in the data, the function
emits a **warning** rather than an error. This is intentional: some callers
may legitimately omit optional columns, and a warning lets you decide whether
the missing column is a problem without crashing the pipeline.

```{r rename-warning, eval=FALSE}
# If no column contains "wait", you get:
#> Warning in mysterycall_rename_columns():
#>   Target substring 'wait' did not match any column name.
#>   Available columns: npi, provider_last_name_legal_name, appointment_offered, ...
#>   Proceeding without renaming for this target.
```

## 5.3 `mysterycall_clean_phase2()`: the full Phase 2 cleaning step

```{r phase2-clean, eval=FALSE}
phase2_result <- mysterycall_clean_phase2(
  data             = "/study/returned_workbooks",  # directory of Excel files
  output_directory = "/study/phase2_output",
  target_strings   = c("appt", "wait", "insurance", "outcome", "caller"),
  new_names        = c("appointment_offered", "wait_days",
                       "insurance_type", "call_outcome", "caller_name"),
  verbose          = TRUE
)
```

When `data` is a **directory path**, `mysterycall_clean_phase2()` reads
every `.xlsx` and `.xls` file in that directory, applies
`mysterycall_rename_columns()` to each one individually, then row-binds them
into a single data frame. This is the recommended approach for the standard
workflow.

When `data` is already a **data frame**, it is processed directly. This is
useful when you have already combined the workbooks manually or want to
process a single file.

A complete example showing messy column normalisation:

```{r phase2-full, eval=FALSE}
# Simulate two messy returned workbooks
alice_wb <- data.frame(
  NPI          = c("1234567893", "9876543210"),
  `Last Name`  = c("Smith", "Jones"),
  `Appt Offered?` = c("Yes", "No"),
  `Wait (days)` = c(14L, NA_integer_),
  check.names  = FALSE
)

bob_wb <- data.frame(
  npi            = c("1122334455", "5544332211"),
  provider_lname = c("Garcia", "Patel"),
  appointment_offered = c("YES", "offered"),
  wait_days           = c(7L, 21L)
)

# Phase 2 standardises both:
# After mysterycall_rename_columns() each workbook will have:
#   npi, provider_last_name_legal_name, appointment_offered, wait_days
```

---

# 6. Quality Check Table: `mysterycall_save_quality_table()`

During and immediately after the calling period, study coordinators need a
quick, printable summary showing whether each caller is on track. The quality
check table provides this.

## 6.1 What the table contains

```{r quality-params, eval=FALSE}
mysterycall_save_quality_table(
  data        = phase2_result,
  output_path = "/study/qa/quality_table.csv",
  strata_col  = "insurance_type"   # optional: stratify rows by payer
)
```

The output is a CSV file with one row per caller (and optionally per
insurance stratum) containing:

| Column | Description |
|---|---|
| `caller_name` | Lab assistant name |
| `insurance_type` | Insurance stratum (if `strata_col` supplied) |
| `n_assigned` | Total rows assigned to this caller |
| `n_completed` | Rows where `call_outcome` is non-missing |
| `completion_pct` | `n_completed / n_assigned × 100` |
| `missing_phone_pct` | Rows with no phone number recorded |
| `missing_outcome_pct` | Rows where call outcome was left blank |

## 6.2 Flagging underperforming callers

The standard threshold in most mystery-caller protocols is 80% completion.
Read the table back and flag callers below that threshold:

```{r quality-flag, eval=FALSE}
qa <- read.csv("/study/qa/quality_table.csv")

# Flag callers below 80% completion
qa$flag <- ifelse(qa$completion_pct < 80, "REVIEW", "OK")

# Print callers needing review
knitr::kable(
  qa[qa$flag == "REVIEW", ],
  caption = "Callers below 80% completion threshold"
)
```

Use this table in your weekly coordinator check-ins to identify callers who
may need retraining or whose workbooks may have been corrupted during return.

---

# 7. Coverage Monitoring: `mysterycall_not_contacted_states()`

After field work concludes, verify that the study has achieved geographic
coverage across all intended states. `mysterycall_not_contacted_states()`
identifies states with zero completed calls.

## 7.1 The `all_states` parameter

```{r coverage-params, eval=FALSE}
not_contacted <- mysterycall_not_contacted_states(
  data       = phase2_result,
  all_states = state.abb           # all 50 states, or a custom list
)
```

`all_states` is the **complete expected set** of states for your study. If
your sampling frame includes only the 12 most populous states, pass only
those 12 codes. The function computes the set difference between `all_states`
and the states present in `data` where `call_outcome` is non-missing.

## 7.2 Visualising contacted vs not-contacted states

```{r coverage-table, eval=FALSE}
# Quick cross-tabulation of states by contact status
contacted_states <- unique(phase2_result$state[
  !is.na(phase2_result$call_outcome)
])

status_vec <- setNames(
  ifelse(state.abb %in% contacted_states, "Contacted", "Not contacted"),
  state.abb
)

print(table(status_vec))
#>
#>  Contacted Not contacted
#>         43             7
```

## 7.3 Filtering for follow-up prioritisation

```{r coverage-filter, eval=FALSE}
# Which specific states need follow-up?
cat("States with zero completed calls:\n")
cat(paste(not_contacted$state, collapse = ", "), "\n")
#> States with zero completed calls:
#> AK, MT, ND, SD, VT, WV, WY

# Subset the roster to those states for a targeted second calling wave
followup_roster <- phase2_result[
  phase2_result$state %in% not_contacted$state, ]

cat(nrow(followup_roster), "providers in under-contacted states\n")
```

Geographic gaps discovered at this stage can inform a second wave of calls or
a discussion with your team about whether certain rural states should be
excluded from the denominator with appropriate justification.

---

# 8. One-Call Orchestration: `mysterycall_run_workflow()`

For standard studies where the default pipeline is appropriate,
`mysterycall_run_workflow()` executes all steps in the correct order and
returns a named list of all intermediate and final artifacts.

## 8.1 Full parameter table

| Parameter | Type | Default | Description |
|---|---|---|---|
| `taxonomy_terms` | `character` | required | NUCC taxonomy codes for NPI search |
| `name_data` | `data.frame` | required | Provider name crosswalk (for name standardisation) |
| `phase1_data` | `data.frame` or path | required | Raw roster for Phase 1 |
| `lab_assistant_names` | `character` | required | Caller names for workbook splitting |
| `output_directory` | `character` | required | Root output directory (Phase 1 and workbooks) |
| `phase2_data` | `data.frame` or path | required | Returned workbooks for Phase 2 |
| `phase2_output_directory` | `character` | required | Where Phase 2 output is written |
| `quality_check_path` | `character` | required | Where the QA CSV is written |
| `phase1_output_directory` | `character` | `output_directory` | Override for Phase 1 output |
| `all_states` | `character` | `state.abb` | States in scope for coverage check |
| `verbose` | `logical` | `TRUE` | Print progress to console |

## 8.2 The returned list structure

```{r workflow-list, eval=FALSE}
names(result)
#>  [1] "phase1_result"           "validated_npi"
#>  [3] "workbook_paths"          "phase2_result"
#>  [5] "quality_table"           "not_contacted_states"
#>  [7] "workflow_summary"        "cohort_hash"
#>  [9] "audit_trail_path"        "run_timestamp"
```

Each element is the direct output of the corresponding pipeline function, so
you can pass `result$phase1_result` to downstream analysis functions just as
you would if you had called `mysterycall_clean_phase1()` directly.

## 8.3 A complete realistic call

```{r workflow-full, eval=FALSE}
result <- mysterycall_run_workflow(
  taxonomy_terms      = c(
    "207V00000X",   # Obstetrics & Gynecology
    "207VB0002X",   # Urogynecology
    "207VX0201X"    # Gynecologic Oncology
  ),
  name_data           = name_crosswalk_df,
  phase1_data         = raw_nppes_export,
  lab_assistant_names = c("Alice Nguyen", "Bob Patel",
                          "Carol Smith", "Dana Lee", "Eve Gonzalez"),
  output_directory    = "/network/obgyn_study/phase1",
  phase2_data         = "/network/obgyn_study/returned_workbooks",
  phase2_output_directory = "/network/obgyn_study/phase2",
  quality_check_path  = "/network/obgyn_study/qa/quality_check.csv",
  all_states          = state.abb,
  verbose             = TRUE
)
```

## 8.4 `mysterycall_run_workflow_logged()`: structured JSON logging

For large studies (> 1000 providers) or multi-site studies where you cannot
be present at the console, use `mysterycall_run_workflow_logged()`. It is
**functionally identical** to `mysterycall_run_workflow()` but wraps every
pipeline stage in structured JSON logging:

```{r workflow-logged, eval=FALSE}
result <- mysterycall_run_workflow_logged(
  # ... same parameters as mysterycall_run_workflow() ...
  log_file = "/network/obgyn_study/logs/workflow_run.jsonl"
)
```

Each line of the `.jsonl` log file contains:

```json
{"stage": "phase1_clean", "status": "started",  "timestamp": "2024-03-15T09:22:41Z", "input_rows": 847}
{"stage": "phase1_clean", "status": "completed", "timestamp": "2024-03-15T09:22:43Z", "output_rows": 1694, "duration_sec": 2.1}
{"stage": "npi_validate", "status": "started",  "timestamp": "2024-03-15T09:22:43Z"}
...
```

If the workflow crashes mid-run, the log file shows exactly which stage
failed and what the input state was, so you can restart from that stage
without re-running earlier steps.

---

# 9. Interpreting the Workflow Summary

After `mysterycall_run_workflow()` returns, the first thing to inspect is
`result$workflow_summary`. This is a data frame with one row per pipeline
stage:

```{r workflow-summary, eval=FALSE}
knitr::kable(
  result$workflow_summary,
  caption = "Row counts through the pipeline"
)
```

| Stage | Input rows | Output rows | Rows retained |
|---|---|---|---|
| `phase1_clean` | 847 | 1694 | 100.0% |
| `npi_validate` | 1694 | 1694 | 100.0% |
| `split_workbooks` | 1694 | 1694 | 100.0% |
| `phase2_clean` | 1694 | 1688 | 99.6% |
| `quality_check` | 1688 | — | — |
| `coverage_monitor` | 1688 | — | — |

The 0.4% row drop between split_workbooks and phase2_clean (6 rows) is
typical: it reflects rows that callers deleted from their workbooks, possibly
because they realised a phone number was a fax line or the provider had
retired. Document these exclusions in your CONSORT flowchart.

**Before proceeding to analysis, confirm:**

1. `rows_retained` from Phase 1 is 100% (no rows were silently dropped).
2. NPI validity rate is acceptable (> 95% is typical for NPPES data).
3. Phase 2 completion is consistent with your calling protocol's expected
   non-contact rate.
4. The `cohort_hash` in `result$cohort_hash` matches the hash stored in
   your lab notebook entry for this wave of data collection.

```{r workflow-check, eval=FALSE}
# Programmatic pre-analysis checks
stopifnot(
  result$workflow_summary$rows_retained[1] == 100,
  mean(result$validated_npi$npi_valid, na.rm = TRUE) > 0.95,
  result$cohort_hash == "a3f8c1d2..."   # paste the hash from your lab notebook
)
```

---

# 10. Practical Guidance

## 10.1 Always run the preflight check first

Before passing any data to the workflow, run:

```{r preflight, eval=FALSE}
mysterycall_preflight_check(phase1_data = raw_roster)
```

This validates that all required columns are present, checks for completely
empty data frames, and warns about potential encoding problems in provider
names (accented characters, non-ASCII punctuation) that could cause silent
mismatches later in the pipeline. Addressing preflight warnings costs minutes;
discovering them after six weeks of field work costs months.

## 10.2 Use a network drive for `output_directory`

All team members need read access to the Phase 1 output, the split workbooks,
and the Phase 2 output. Placing `output_directory` on a shared network drive
or a cloud-synced folder (Box, Dropbox, OneDrive) ensures that callers can
download their workbooks without emailing files and that coordinators can
monitor progress in real time.

Avoid paths with spaces. Use hyphens or underscores:

```r
# Good
output_directory = "/Volumes/ResearchDrive/obgyn_mystery_caller_2024"

# Risky on some OS + R version combinations
output_directory = "/Volumes/Research Drive/OBGYN Mystery Caller 2024"
```

## 10.3 Store the `cohort_hash` in your lab notebook

The `cohort_hash` is the single most important artifact for linking your
raw data to your published analysis. Write it in your lab notebook or your
IRB protocol amendment. A typical entry looks like:

> *Wave 1 data collection completed 2024-03-22. Phase 1 cohort hash:
> `a3f8c1d2b5e7f091`. Audit trail at
> `/network/obgyn_study/phase1/audit_a3f8c1d2.json`.*

This record lets you — or a journal auditor — reproduce the exact dataset
used in your analysis three years after the study is complete.

## 10.4 For large studies, use the logged variant

For studies with more than 1,000 providers, use
`mysterycall_run_workflow_logged()`. If Phase 2 cleaning fails on a
corrupted workbook at 2 AM, the log file tells you exactly which workbook
caused the failure. You can fix that one file and re-run Phase 2 alone rather
than restarting the entire workflow.

The logged variant also produces timing information for each stage. In large
studies, Phase 1 with `duplicate_rows = TRUE` and `output_format = "parquet"`
is typically the fastest combination (5–10× faster than CSV for > 5,000 rows).

```{r large-study, eval=FALSE}
# Recommended settings for large studies
result <- mysterycall_run_workflow_logged(
  phase1_data      = large_roster,
  output_format    = "parquet",   # faster than CSV for > 5,000 rows
  duplicate_rows   = TRUE,
  verbose          = FALSE,       # suppress console output in batch scripts
  log_file         = "/logs/workflow.jsonl"
)
```

---

# 11. Business-Day Wait Times

Mystery-caller wait times are measured in **business days** — weekends and US
federal holidays are excluded. `mysterycall_count_business_days()` counts
business days between two dates; `mysterycall_business_days()` applies the
calculation to every row of a data frame.

```{r business-days-scalar, echo=TRUE, eval=requireNamespace("bizdays", quietly = TRUE)}
# How many business days between a call on a Monday and an appointment on
# Friday of the following week?  (Calendar days = 11; business days = 10)
bd <- mysterycall_count_business_days(
  start_date = "2024-03-04",   # Monday
  end_date   = "2024-03-15"    # Friday two weeks later
)
cat("Business days:", bd, "\n")
```

```{r business-days-df, echo=TRUE, eval=requireNamespace("bizdays", quietly = TRUE)}
# Apply to a data frame of completed calls
calls_df <- data.frame(
  npi       = paste0("NPI-", 1:6),
  insurance = c("Medicaid", "Private", "Medicaid", "Private", "Medicare", "Uninsured"),
  call_date = as.Date(c("2024-03-04", "2024-03-04", "2024-03-11",
                         "2024-03-11", "2024-03-18", "2024-03-18")),
  appt_date = as.Date(c("2024-03-25", "2024-03-12", "2024-04-01",
                         "2024-03-19", "2024-03-27", "2024-04-08"))
)

# mysterycall_business_days() column params: call_col, appt_col, result_col
calls_with_bd <- mysterycall_business_days(
  data       = calls_df,
  call_col   = "call_date",
  appt_col   = "appt_date",
  result_col = "wait_days"
)

knitr::kable(
  calls_with_bd[, c("npi", "insurance", "call_date", "appt_date", "wait_days")],
  col.names = c("NPI", "Insurance", "Call date", "Appt date", "Wait (business days)"),
  caption   = "Business-day wait times computed by `mysterycall_business_days()`."
)
```

---

# 12. Caller Quality Control

After data collection is complete, audit each caller's performance before
proceeding to analysis.

## 12.1 Inter-rater reliability

`mysterycall_caller_reliability()` computes Cohen's kappa (for categorical
outcomes) or the intraclass correlation coefficient (for numeric outcomes) on
records audited by two callers.

```{r caller-reliability, echo=TRUE, eval=TRUE}
set.seed(7)
# Long-format audit sample: each record appears twice — once per caller
n_records <- 30L
audit_long <- data.frame(
  record_id  = rep(1:n_records, 2),
  caller     = rep(c("Alice", "Bob"), each = n_records),
  appt_offered = c(
    sample(c("Yes", "No"), n_records, replace = TRUE, prob = c(0.72, 0.28)),
    sample(c("Yes", "No"), n_records, replace = TRUE, prob = c(0.68, 0.32))
  )
)

rel <- mysterycall_caller_reliability(
  data        = audit_long,
  caller_col  = "caller",
  outcome_col = "appt_offered",
  type        = "kappa"
)

print(rel)
```

## 12.2 Per-caller productivity

`mysterycall_call_productivity()` tallies completed, pending, and failed calls
by caller and returns a summary data frame for QC review.

```{r call-productivity, echo=TRUE, eval=TRUE}
set.seed(8)
n_rows <- 80L
prod_df <- data.frame(
  caller           = sample(c("Alice", "Bob", "Carol", "Dana"),
                             n_rows, replace = TRUE),
  call_date        = sample(seq(as.Date("2024-02-01"), as.Date("2024-03-31"),
                                 by = "day"), n_rows, replace = TRUE),
  appointment_offered = sample(c(1L, 1L, 1L, 0L), n_rows, replace = TRUE)
)

prod <- mysterycall_call_productivity(
  data        = prod_df,
  caller_col  = "caller",
  date_col    = "call_date",
  outcome_col = "appointment_offered"
)

knitr::kable(
  prod$summary,
  caption = "Per-caller productivity from `mysterycall_call_productivity()`."
)
```
