---
title: "Writing the Results Section"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
    df_print: kable
description: >
  Assemble a ready-to-paste manuscript Results section from mysterycall's prose
  builders, with every direction word tied to the sign of the data so the
  narrative can never contradict the tables.
vignette: >
  %\VignetteIndexEntry{Writing the Results Section}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(mysterycall)
```

## Why generate Results prose programmatically?

The Results section of an access-disparities manuscript restates numbers that
already live in your model output and tables. Re-typing them by hand invites a
whole class of bug: the sentence says "increasing" while the coefficient is
negative, an odds ratio below 1 is described as "more likely", or a confidence
interval that crosses the null is written as if it were bounded away from zero.

`mysterycall` provides a family of **prose builders** that take your estimates
and emit manuscript-ready sentences whose wording is derived from the *sign and
magnitude of the data*. Wire the narrative to these builders and the words can
never drift out of step with the tables.

| Builder | Input | Produces |
|---|---|---|
| `mysterycall_results_paragraph()` | OR table (logistic) | odds-of-acceptance sentences |
| `mysterycall_write_results_paragraph()` | IRR table (Poisson/NB) | wait-time IRR sentences |
| `mysterycall_irr_to_days()` | IRR table + reference mean | absolute day differences (signed CI) |
| `mysterycall_wait_time_sentence()` | raw data | median/IQR + log-rank-style summary |
| `mysterycall_get_direction()` | any signed number | "higher"/"lower" (configurable) |
| `mysterycall_get_change_verb()` | any signed number | "increasing"/"decreasing" (configurable) |

Every builder here works on a plain data frame, so this vignette needs no model
fitting — but the same objects are returned by `mysterycall_logistic_model()`,
`mysterycall_poisson_model()`, and `mysterycall_nb_model()`.

## A worked example

Two physician subspecialties are compared with a reference group for (a) whether
a new-patient appointment was offered and (b) how long the wait was. We build the
estimate tables directly; in practice they come from your fitted models'
`$or_table` / `$irr_table` elements.

```{r tables}
# Odds ratios for "appointment offered" (logistic regression output)
or_tbl <- data.frame(
  term     = c("(Intercept)", "insuranceMedicaid", "insuranceMedicare"),
  or       = c(2.10, 0.62, 0.88),
  ci_lower = c(1.20, 0.41, 0.63),
  ci_upper = c(3.68, 0.94, 1.23),
  p_value  = c(0.008, 0.024, 0.451),
  stringsAsFactors = FALSE
)

# Incidence rate ratios for wait days (Poisson/NB regression output).
# `p_value` (numeric) is used by write_results_paragraph(); `p_value_fmt`
# (character) is used by irr_to_days() for the in-sentence p-value.
irr_tbl <- data.frame(
  term        = c("(Intercept)", "insuranceMedicaid", "insuranceMedicare"),
  irr         = c(1.00, 1.28, 1.04),
  ci_lower    = c(NA,   1.05, 0.90),
  ci_upper    = c(NA,   1.56, 1.20),
  p_value     = c(NA,   0.014, 0.590),
  p_value_fmt = c(NA,   "0.014", "0.590"),
  stringsAsFactors = FALSE
)
```

### Appointment acceptance (odds ratios)

```{r acceptance}
acceptance <- mysterycall_results_paragraph(
  or_tbl,
  ref_group     = "commercial insurance",
  exposure_col  = "insurance",
  outcome_label = "a new-patient appointment"
)
cat(acceptance)
```

The intro clause claims significance **only** when a level actually clears
`alpha`; the direction ("less likely" / "more likely" / "similar odds") is chosen
from whether each OR sits below, above, or at 1.

### Wait-time incidence rate ratios

```{r wait-irr}
wait_irr <- mysterycall_write_results_paragraph(
  irr_tbl,
  ref_group     = "commercial insurance",
  exposure_col  = "insurance",
  outcome_label = "time to appointment"
)
cat(wait_irr)
```

### Translating IRRs to absolute days (signed CIs)

IRRs are hard for readers to feel. `mysterycall_irr_to_days()` multiplies them by
the reference-group mean wait to report absolute day differences — and it keeps
the **sign** of the confidence interval, flagging any interval that crosses zero
as not statistically significant.

```{r irr-days}
days <- mysterycall_irr_to_days(
  irr_tbl,
  baseline_mean = 21,          # mean wait (days) in the reference group
  exposure_col  = "insurance",
  ref_group     = "commercial insurance"
)
cat(days$paragraph)
```

Note the Medicare row: because its day-scale CI spans zero, the sentence says so
rather than implying a real effect.

For a non-insurance exposure (e.g. subspecialty), drop the insurance-specific
wording with `exposure_descriptor = NULL`:

```{r irr-days-ent}
ent_tbl <- data.frame(
  term        = c("(Intercept)", "ent_typeLaryngology"),
  irr         = c(1.00, 1.20),
  ci_lower    = c(NA,   1.02),
  ci_upper    = c(NA,   1.41),
  p_value_fmt = c(NA,   "0.028"),
  stringsAsFactors = FALSE
)
ent_days <- mysterycall_irr_to_days(
  ent_tbl, baseline_mean = 23, exposure_col = "ent_type",
  ref_group = "General", subject = "callers", exposure_descriptor = NULL
)
cat(ent_days$sentences[1])
```

### Descriptive wait-time sentence

Straight from the raw data, no model required:

```{r wait-sentence}
set.seed(1)
raw <- data.frame(
  wait_days = c(rpois(60, 12), rpois(60, 16)),
  insurance = rep(c("Private", "Medicaid"), each = 60)
)
wt <- mysterycall_wait_time_sentence(
  raw, outcome_col = "wait_days", group_col = "insurance",
  reference = "Private"
)
cat(wt$sentence)
```

### Direction words tied to the data sign

For any custom sentence — a change over time, a difference between groups — let
the builder choose the word from the sign so the prose cannot contradict the
number. This is exactly the tool for subspecialty/workforce framing.

```{r direction}
# A projected annual change in provider supply
slope <- -0.8
sprintf(
  "The maternal-fetal-medicine workforce is %s (%+.1f providers per year).",
  mysterycall_get_change_verb(slope), slope
)

# A between-group wait difference, with custom vocabulary
diff_days <- 4.3
sprintf(
  "Medicaid callers waited %s than commercial-insurance callers (%.1f days).",
  mysterycall_get_direction(diff_days, positive = "longer", negative = "shorter"),
  abs(diff_days)
)
```

Both are vectorized and `NA`-safe, and a `tol` argument defines a no-change band
around zero (e.g. `mysterycall_get_change_verb(x, tol = 0.5)` calls anything
within ±0.5 `"stable"`).

## Assembling the full Results section

Concatenate the pieces into one narrative. Because every clause was generated
from the estimates, editing a number in the model re-flows the prose correctly.

```{r assemble}
results_section <- paste(
  acceptance,
  wait_irr,
  days$paragraph,
  sep = "\n\n"
)
cat(results_section)
```

## Model-diagnostic prose

Two more builders describe the model itself. They take a *fitted* model object,
so they are shown here without evaluation:

```{r diagnostics, eval = FALSE}
# Variance explained (marginal / conditional R-squared) for a mixed model
cat(mysterycall_r2_sentence(fit))

# Dispersion check for a count model
cat(mysterycall_overdispersion_sentence(fit))
```

## Best practice

- **Never hand-type a direction word.** Feed the estimate to
  `mysterycall_get_direction()` / `mysterycall_get_change_verb()` (or use a
  builder that does). A sentence that says "increasing" over a negative slope is
  a data-integrity bug, not a typo.
- **Report absolute effects alongside ratios.** `mysterycall_irr_to_days()`
  gives readers days, with signed CIs that honestly flag null results.
- **Regenerate, don't edit.** Re-run the builders after any model change so the
  Results text and the tables always agree.

```{r session, echo = FALSE}
sessionInfo()
```
